---
title: "Model Assumptions and Diagnostics"
execute:
  echo: true
  warning: false
  message: false
format: 
  revealjs:
    theme: default
    self-contained: true
    slide-number: true
    width: 1600
    height: 900
    df-print: paged
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

## Introduction


```{r setup}
#| echo: false
library(tidyverse)
library(ssstats)
library(broom)
```

- Recall the glm, 

$$ 
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + ... \beta_k x_k + \varepsilon 
$$ 

- In this model, $\varepsilon$ is the *residual error* term.

- Recall that the residual error (how far away the observation is from the response surface) is defined as 

$$\varepsilon = y - \hat{y}$$ 

- Note that

    - $y$ is the observed value
    
    - $\hat{y}$ is the predicted value

## Model Assumptions: HILE+G

- To ensure the validity of the least squares estimation, we have several assumptions:

    - Homogeneity of variance

    - Independence of observations

    - Linearity

    - Existence

    - Gaussian errors
    
- Think: HILE+G

    - We require HILE for the OLS to be valid.
    - We require +G for maximum likelihood estimation.

## Model Assumptions: Homogeneity of Variance

- Homogeneity of variance (also called homoscedasticity) means that the variance of the residuals is constant across all levels of the predictors.

- If the variance of the residuals changes at different levels of the predictors, we have heteroscedasticity.

- We will check this assumption graphically using a scatterplot of residual vs. fitted values.

    - We do not want to see a pattern.
    
    - A "funnel" or "open fan" shape indicates heteroscedasticity.
    
    - A general "cloud" shape indicates homoscedasticity.
    
## Model Assumptions: Homogeneity of Variance (R)

- We will use the `ssstats` package to check assumptions.

- For variance, we can use the `variance_check()` function.

```{r}
#| eval: false
model %>% variance_check()
```

## Model Assumptions: Homogeneity of Variance

- Suppose we have a model, `m1`,

$$
\hat{y} = 3 + 1.5 x + \hat{\varepsilon}
$$

```{r}
#| echo: false
set.seed(9176)
reg_data <- tibble(id = 1:100,
                   x  = runif(100, 0, 10),
                   epsilon = rnorm(100, mean = 0, sd = 2)) %>%
  mutate(y = 3 + 1.5 * x + epsilon)

m1 <- glm(y ~ x, data = reg_data)
```


```{r}
#| fig-align: 'center'
m1 %>% variance_check()
```

## Model Assumptions: Homogeneity of Variance

- Suppose we have a model, `m2`,

$$
\hat{y} = 4 + 1.2 x_1 - 0.8 x_2 + \hat{\varepsilon}
$$

```{r}
#| echo: false
set.seed(651965)
reg_data <- tibble(
  id = 1:150,
  x1 = runif(150, 0, 10),   # predictor 1
  x2 = runif(150, 0, 5),    # predictor 2
  epsilon = rnorm(150, mean = 0, sd = 2)  # homoskedastic errors
) %>%
  mutate(
    y = 4 + 1.2 * x1 - 0.8 * x2 + epsilon
  )
m2 <- glm(y ~ x1 + x2, data = reg_data)
```


```{r}
#| fig-align: 'center'
m2 %>% variance_check()
```

## Model Assumptions: Homogeneity of Variance

- Suppose we have a model, `m3`,

$$
\hat{y} = 5 + 1.1 x_1 - 0.9 x_2 + \hat{\varepsilon}
$$

```{r}
#| echo: false
set.seed(19877)
n <- 150
reg_data <- tibble(
  id = 1:n,
  x1 = runif(n, 0, 10),
  x2 = runif(n, 0, 5),
  # Non-normal but equal-variance errors
  epsilon = rexp(n, rate = 1) - 1   # mean-centered exponential
) %>%
  mutate(
    y = 5 + 1.1 * x1 - 0.9 * x2 + epsilon
  )
m3 <- glm(y ~ x1 + x2, data = reg_data)
```

```{r}
#| fig-align: 'center'
m3 %>% variance_check()
```

## Model Assumptions: Homogeneity of Variance

- Suppose we have a model, `m4`,

$$
\hat{y} = 4 + 1.0 x_1 - 0.7 x_2 + \hat{\varepsilon}
$$

```{r}
#| echo: false
set.seed(01759)
n <- 150
reg_data <- tibble(
  id = 1:n,
  x1 = runif(n, 0, 10),
  x2 = runif(n, 0, 5)
) %>%
  mutate(
    # Variance increases with x1
    sigma = 0.5 + 0.4 * x1,
    # Normal errors, but heteroskedastic
    epsilon = rnorm(n, mean = 0, sd = sigma),
    y = 4 + 1.0 * x1 - 0.7 * x2 + epsilon
  )
m4 <- glm(y ~ x1 + x2, data = reg_data)
```

```{r}
#| fig-align: 'center'
m4 %>% variance_check()
```

## Model Assumptions: Homogeneity of Variance

- Suppose we have a model, `m5`,

$$
\hat{y} = 3 + 1.3 x_1 - 0.9 x_2 + \hat{\varepsilon}
$$
    
```{r}
#| echo: false
set.seed(5715)
n <- 200
reg_data <- tibble(
  id = 1:n,
  x1 = runif(n, 0, 10),
  x2 = runif(n, 0, 5)
) %>%
  mutate(
    # Strong variance inflation → obvious fan
    sigma = 0.3 + 0.8 * x1,
    # Normal but heteroskedastic errors
    epsilon = rnorm(n, mean = 0, sd = sigma),
    y = 3 + 1.3 * x1 - 0.9 * x2 + epsilon
  )
m5 <- glm(y ~ x1 + x2, data = reg_data)
```
    
```{r}
#| fig-align: 'center'
m5 %>% variance_check()
```    

## Model Assumptions: Homogeneity of Variance

- Suppose we have a model, `m6`,

$$
\hat{y} = 2 + 1.1 x_1 - 0.8 x_2 + \hat{\varepsilon}
$$

```{r}
#| echo: false
set.seed(855)
n <- 250
wild_data <- tibble(
  id = 1:n,
  x1 = runif(n, 0, 10),
  x2 = runif(n, 0, 5)
) %>%
  mutate(
    # variance explodes with x1 (big-time fan opening)
    sigma = 0.2 + 0.25 * (x1^2),
    # non-normal core: heavy tails (t-dist with low df)
    eps_core = rt(n, df = 2) * sigma,
    # add skew: exponential shock (mean-centered) scaled by sigma
    eps_skew = (rexp(n, rate = 1) - 1) * (0.6 * sigma),
    # occasional "catastrophic" outliers (mixture component)
    is_outlier = rbinom(n, size = 1, prob = 0.08),
    eps_outlier = is_outlier * rnorm(n, mean = 0, sd = 10 * sigma),
    epsilon = eps_core + eps_skew + eps_outlier,
    # response
    y = 2 + 1.1 * x1 - 0.8 * x2 + epsilon
  ) %>%
  select(id, x1, x2, y)
m6 <- glm(y ~ x1 + x2, data = wild_data)
```

```{r}
#| fig-align: 'center'
m6 %>% variance_check()
```   

## Model Assumptions: Independence of Observations

- We also assume that our observations are *independent*.

    - This means that the value of one observation does not influence or provide information about another observation.
    
- Examples of *independent data*:

    - Randomly sampled individuals from a large population.
    - Measurements taken from different individuals with no inherent relationship         

- Examples of *dependent data*:
    
    - Measurements taken from the same individual over time (repeated measures)
    - Measurements taken from individuals within the same group or cluster (e.g., students within the same classroom; people in the same family)
    - Spatial data where observations are collected from nearby locations.
        
## Model Assumptions: Independence of Observations   

- Violations of independence often occur in longitudinal, time series, and spatial data.

- When analyzing correlated data, the methodology learned in this course *is not appropriate*.

    - Honestly, it's really not a big deal: we include a covariance structure in the model to account for dependence.
    
    - However, this methodology is *beyond the scope* of this course -- we will focus only on independent data.
    
- What about squicky cases?

    - We know that perfect independence may not be achievable in practice.
    
    - Mild violations of independence may not severely impact results, but this is context-dependent.
    
    - Ultimately, we do the best with what we can and notate any potential limitations in our analysis.
    
## Model Assumptions: Linearity    

- We assume that the relationship between each predictor and the response is linear in the *parameters* (i.e., $\beta_i$).

    - This means that the effect of a one-unit change in a predictor on the response is constant, regardless of the value of that predictor.
    
- In this course, all of the models we will construct are *linear models*. 

- The language gets tricky because we can have non-linear relationships in a linear model.

    - For example, we can include polynomial terms (e.g., $x^2$, $x^3$) or interaction terms (e.g., $x_1 \times x_2$) in a linear model.
    
    - As long as the model is linear <u>in the parameters</u>, it is considered a linear model.
    
## Model Assumptions: Linearity 

- Consider the following models -- which are linear in $\beta$?

    1. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \hat{\varepsilon}$
    
    2. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2 x_1^2 + \hat{\varepsilon}$
    
    3. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2 x_2 x_3 + \hat{\varepsilon}$
    
    4. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2^2 x_2 + \hat{\varepsilon}$
    
    5. $\hat{y} = \beta_0 + \beta_1 \log(x_1) + \beta_2 x_2 + \hat{\varepsilon}$
    
    6. $\hat{y} = \beta_0 + \log(\beta_1) x_1 + \beta_2 x_2 + \hat{\varepsilon}$
    
## Model Assumptions: Linearity 

- Consider the following models -- which are linear in $\beta$?

    1. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \hat{\varepsilon}$ -- linear
    
    2. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2 x_1^2 + \hat{\varepsilon}$ -- linear
    
    3. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2 x_2 x_3 + \hat{\varepsilon}$ -- linear
    
    4. $\hat{y} = \beta_0 + \beta_1 x_1 + \beta_2^2 x_2 + \hat{\varepsilon}$ -- broken by $\beta_2^2$
    
    5. $\hat{y} = \beta_0 + \beta_1 \log(x_1) + \beta_2 x_2 + \hat{\varepsilon}$ -- linear
    
    6. $\hat{y} = \beta_0 + \log(\beta_1) x_1 + \beta_2 x_2 + \hat{\varepsilon}$ -- broken by $\log(\beta_1)$
    
## Model Assumptions: Existence

- The residuals have a finite mean and variance.

    - Mean (first moment): $E[X] = \mu$
    - Variance (second central moment): $E[(X - \mu)^2] = \sigma^2$

- This assumption is almost always satisfied in practical applications.

    - We can think of this as: the residuals are not so extreme that their average or variability is infinite.
    
- This is not an assumption we check -- we will see estimation errors when this assumption is violated.

## Model Assumptions: Gaussian Distribution

- Note! The normal distribution is also known as the Gaussian distribution.

- For our inference to be valid, we must assume that the residuals are normally distributed.

    - Otherwise, our estimation may not be accurate (i.e., inference may lead us to the "wrong" decision).

- We will check this assumption using a quantile-quantile (q-q) plot of the residuals.

    - We want to see points roughly along the line.
    
    - Large deviations from the line indicate non-normality.
    
- We can also use a histogram to "back up" our decision from the q-q plot.   

    - We want to see it roughly mound-shaped and symmetric.
    
    - Skewed or multi-modal histograms indicate non-normality.
    
## Model Assumptions: Gaussian Distribution (R)

- We will use the `ssstats` package to check assumptions.

- For normality, we can use the `normality_check()` function.

```{r}
#| eval: false
model %>% normality_check()
```

## Model Assumptions: Gaussian Distribution 

- Recall our example model, `m1`:

$$
\hat{y} = 3 + 1.5 x + \hat{\varepsilon}
$$

```{r}
#| fig-align: 'center'
m1 %>% normality_check()
```


## Model Assumptions: Gaussian Distribution 

- Recall our example model, `m2`:

$$
\hat{y} = 4 + 1.2 x_1 - 0.8 x_2 + \hat{\varepsilon}
$$

```{r}
#| fig-align: 'center'
m2 %>% normality_check()
```

## Model Assumptions: Gaussian Distribution 

- Recall our example model, `m3`:

$$
\hat{y} = 5 + 1.1 x_1 - 0.9 x_2 + \hat{\varepsilon}
$$

```{r}
#| fig-align: 'center'
m3 %>% normality_check()
```

## Model Assumptions: Gaussian Distribution 

- Recall our example model, `m4`:

$$
\hat{y} = 4 + 1.0 x_1 - 0.7 x_2 + \hat{\varepsilon}
$$

```{r}
#| fig-align: 'center'
m4 %>% normality_check()
```

## Model Assumptions: Gaussian Distribution 

- Recall our example model, `m5`:

$$
\hat{y} = 3 + 1.3 x_1 - 0.9 x_2 + \hat{\varepsilon}
$$

```{r}
#| fig-align: 'center'
m5 %>% normality_check()
```

## Model Assumptions: Gaussian Distribution 

- Recall our example model, `m6`:

$$
\hat{y} = 2 + 1.1 x_1 - 0.8 x_2 + \hat{\varepsilon}
$$

```{r}
#| fig-align: 'center'
m6 %>% normality_check()
```

## Model Assumptions (R)

- To summarize, our model assumptions can be written as:

$$\varepsilon \overset{\text{iid}}{\sim} N(0, \sigma^2)$$

- This notation combines our checkable assumptions:

    - Residuals are independent and identically distributed (I)
    
    - Residuals follow a normal distribution (G), centered at 0 and with some constant variance, $\sigma^2 (H)$.
    
- We can look at a graph with all of these checks at once using the `reg_check()` function in the `ssstats` package.

```{r}
#| eval: false
model %>% reg_check()
```
    
## Model Assumptions

- Consider this for `m1`:

```{r}
#| fig-align: 'center'
m1 %>% reg_check()
```

## Model Assumptions

- Consider this for `m3`:

```{r}
#| fig-align: 'center'
m3 %>% reg_check()
```

## Model Assumptions

- Consider this for `m5`:

```{r}
#| fig-align: 'center'
m5 %>% reg_check()
```


## Testing for Normality...?

- No.

    - Formal tests for normality (e.g., Shapiro-Wilk test, Kolmogorov-Smirnov test) are generally not recommended in practice.
    
    - These tests are sensitive to large sample sizes, leading to the rejection of normality, even for minor infractions that do not affect our inferential conclusions.
    
        - As sample size ($n$) increases, our standard error ($\text{s.d.} / \sqrt{n}$) decreases, increasing our test statistic ($\text{observed} / \text{s.e.}_{\hat{\beta}_i}$), decreasing the $p$-value...... making it easier to reject $H_0$: the data follows a normal distribution.
    
    - Instead, graphical methods (q-q plots, histograms) and practical considerations are preferred for assessing normality.

## Wrap Up

- This lecture has covered the assumptions on the linear model for continuous outcomes assuming independent data and Gaussian errors.

- My approach in "real life" is:

    1. Fit the model.
    
    2. Check assumptions using graphical diagnostics.
    
    3. If assumptions are violated, consider an alternative approach.
    
    4. Document any assumption violations and their potential impact on results.
    
    5. Discuss findings (including relevant failed assumption checks) with collaborators/stakeholders.
    
- Note that this is a simplified overview; real-world data analysis may require more nuanced approaches with field/application-specific considerations and methodologies.