---
title: "Visualization of the Model: <br>Multiple Linear Regression"
execute:
  echo: true
  warning: false
  message: false
format: 
  revealjs:
    theme: default
    self-contained: true
    slide-number: true
    width: 1600
    height: 900
    df-print: paged
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

## Introduction

- Recall the general linear model, $$y = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k$$

    - $\beta_0$ is the $y$-intercept, or the average outcome ($y$) when all $x_i = 0$.

    - $\beta_i$ is the slope for predictor $i$ and describes the relationship between the predictor and the outcome, after adjusting (or accounting) for the other predictors in the model.
    
- In the last lecture, we used a linear model to explore the relationships between clubhouse happiness and laughs, grumbles, and time spent with friends.    

## Lecture Example Set Up

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
using echo: false here is just to make the slides cleaner.
-->

```{r}
#| echo: false
library(tidyverse)
library(broom)
```

- On a busy day at the clubhouse, Mickey Mouse wants to understand what drives "happiness" at the end of the day. For each day, he records (in the <u>**clubhouse**</u> dataset):

    - Time with friends (in hours; *time_with_friends*): how many hours Mickey spends hanging out with his friends.
    - Goofy Laughs (a count; *goofy_laughs*) – how many big goofy laughs happen that day.
    - Donald Grumbles (a count; *donald_grumbles*): how many times Donald gets frustrated and grumbles.
    - Clubhouse Happiness (a score; *clubhouse_happiness*): an overall happiness score at the end of the day.
    
```{r}
clubhouse <- read_csv("~/Desktop/GitHub/samanthaseals/STA4231_STA6262/files/data/lectures/W1_mickey_clubhouse.csv")
```    



## Example 1: Simple Linear Regression

- For our first example, let's look at clubhouse happiness as a function of time spent with friends.

```{r}
m1 <- glm(clubhouse_happiness ~ time_with_friends, 
         family = "gaussian",
         data = clubhouse)
m1 %>% tidy()
```

$$
\hat{\text{happiness}} = 57.55 + 3.49 \text{ time}
$$

## Model Visualization: Simple Linear Regression

- We can visualize this simple linear regression model with a scatterplot and regression line.

- To create the regression line, we need to create predicted values from our model.

```{r}
clubhouse <- clubhouse %>%
  mutate(predicted_happiness_k1 = 57.55 + 3.49*time_with_friends)
```

<!-- 
note: I only include the code below to make a few obs
to the slides. please do not use head() on your assignments
-->

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
using echo: false here is just to make the slides cleaner.
-->

```{r}
#| echo: false
clubhouse %>% 
  select(clubhouse_happiness, time_with_friends, predicted_happiness_k1) %>% 
  head()
```

- Now we can plot the data and the regression line.

## Plotting using `library(ggplot2)`

- The `ggplot()` function initializes a ggplot object.

```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot()
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot()
```

## Plotting using `library(ggplot2)`

```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot()
```

## Plotting using `library(ggplot2)`

- We must define the aesthetics (i.e., the $x$ and $y$ variables) inside `ggplot()`.
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y))
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness))
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness))
```

## Plotting using `library(ggplot2)`

- We must add `geom_TYPE()` layers to actually see anything on the plot.

    - We layer multiple `geom`s to the plot using `+` operator. 
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y)) +
  geom_TYPE()
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point()
```

## Plotting using `library(ggplot2)`

- We continue to layer our plot with additional `geom_TYPE()`s,
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y)) +
  geom_TYPE() +
  geom_TYPE()
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line()
```

## Plotting using `library(ggplot2)`

- Ooops! That `geom_line()` didn't work as expected. 

    - In the `ggplot()`, we set $y$ to be the actual happiness values (*clubhouse_happiness*).
    - We now need to overwrite the $y$ variable to be the the predicted values from our model (*predicted_happiness*).
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y)) +
  geom_TYPE() +
  geom_TYPE(aes(y = predicted_y))
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1))
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1))
```

## Plotting using `library(ggplot2)`

- Now, we can work on "prettying" up our plot.

    - My first step is to change the theme using `theme_NAME()`.
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y)) +
  geom_TYPE() +
  geom_TYPE(aes(y = predicted_y)) +
  theme_NAME()
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1)) +
  theme_bw()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1)) +
  theme_bw()
```

## Plotting using `library(ggplot2)`

- Now, we can work on "prettying" up our plot.

    - Then, I want to clean up the axis titles.
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y)) +
  geom_TYPE() +
  geom_TYPE(aes(y = predicted_y)) +
  labs(x = "x axis title",
       y = "y axis title") +
  theme_NAME()
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1)) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1)) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Plotting using `library(ggplot2)`

- Now, we can work on "prettying" up our plot.

    - We could also add a graph title.
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y)) +
  geom_TYPE() +
  geom_TYPE(aes(y = new_y)) +
  labs(x = "x axis title",
       y = "y axis title",
       title = "title of graph") +
  theme_NAME()
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1)) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness",
       title = "Predicted relationship between happiness and time spent with friends") +
  theme_bw()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point() + 
  geom_line(aes(y = predicted_happiness_k1)) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness",
       title = "Predicted relationship between happiness and time spent with friends") +
  theme_bw()
```

## Plotting using `library(ggplot2)`

- Now, we can work on "prettying" up our plot.

    - Outside of `aes()`, we can specify colors, line types, point shapes, etc.
    
```{r}
#| eval: false
#| fig-align: 'center'
dataset_name %>% ggplot(aes(x = variable_on_x,
                            y = variable_on_y)) +
  geom_TYPE(color = "#HEX", size = size_number) +
  geom_TYPE(aes(y = new_y), color = "#HEX", size = size_number) +
  labs(x = "x axis title",
       y = "y axis title",
       title = "title of graph") +
  theme_NAME()
```

- In our example,

```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_k1), color = "#004C97", size = 1.5) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
(k1 <- clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_k1), color = "#004C97", size = 1.5) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw())
```

## Example 2: Multiple Regression (*k* = 2)

- For our second example, let's look at clubhouse happiness (*clubhouse_happiness*) as a function of time spent with friends (*time_with_friends*) and big, goofy laughs (*goofy_laughs*).

```{r}
m2 <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs, 
          family = "gaussian",
          data = clubhouse)
m2 %>% tidy()
```

$$
\hat{\text{happiness}} = 39.25 + 3.06 \text{ time} + 0.66 \text{ laughs}
$$

## Model Visualization: Multiple Regression

- Now that there's an additional predictor, we can't easily visualize the model with a simple 2D scatterplot.

    - In theory, we could create a 3D scatterplot with a regression plane, but those are hard to read and interpret.
    
- Instead, we will visualize the relationship between $y$ (clubhouse happiness) and $x_1$ (one predictor) while holding $x_2$ (the other predictor) constant.
    
- In our example,

    - We will visualize the relationship between clubhouse happiness and time spent with friends.
    
    - Time spent with friends will be on the $x$-axis and allowed to vary.
    
    - We will hold goofy laughs constant at some value.
    
        - With continuous predictors, I typically plug in the `median()` when drafting initial graphs for collaborators.
        
## Model Visualization: Multiple Regression        

```{r}
clubhouse <- clubhouse %>%
  mutate(predicted_happiness_k2 = 39.25 + 3.06*time_with_friends + 0.66*median(goofy_laughs))
```

<!-- 
note: I only include the code below to make a few obs
to the slides. please do not use head() on your assignments
-->

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
using echo: false here is just to make the slides cleaner.
-->

```{r}
#| echo: false
clubhouse %>% 
  mutate(goofy_laughs = median(goofy_laughs)) %>%
  select(time_with_friends, goofy_laughs) %>% 
  head()
```

## Plotting using `library(ggplot2)`

- Then, constructing our graph,
    
```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_k2), color = "#004C97", size = 1.5) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
(k2 <- clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_k2), color = "#004C97", size = 1.5) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw())
```

## Example 3: Multiple Regression (*k* = 3)

- For our third example, let's return to our full model.

- We looked at clubhouse happiness (*clubhouse_happiness*) as a function of time spent with friends (*time_with_friends*), big, goofy laughs (*goofy_laughs*), and how much Donald grumbles (*donald_grumbles*).

```{r}
m3 <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles,
          family = "gaussian",
          data = clubhouse)
m3 %>% tidy()
```

$$
\hat{\text{happiness}} = 47.58 + 3.58 \text{ time} + 0.66 \text{ laughs} - 1.06 \text{ grumbles}
$$

## Model Visualization: Multiple Regression

- In this example, we have $k=3$ predictors.

    - We can't easily visualize the model with a simple 2D scatterplot or even a 3D scatterplot.
    
- Instead, we will visualize the relationship between $y$ (clubhouse happiness) and $x_1$ (one predictor) while holding all other $x_i$ (the other predictors) constant.

    - One $x_i$ will vary on the $x$-axis.
    - We will plug in plausible values for the other predictors.
    
- In our example,

    - We will visualize the relationship between clubhouse happiness and time spent with friends.
    
    - Time spent with friends will be on the $x$-axis and allowed to vary.
    
    - We will hold goofy laughs constant at some value.
    - We will also hold Donald grumbles constant at some value.

## Model Visualization: Multiple Regression

```{r}
clubhouse <- clubhouse %>%
  mutate(predicted_happiness_k3 = 47.58 + 3.58*time_with_friends + 0.66*median(goofy_laughs) - 1.06*median(donald_grumbles))
```

<!-- 
note: I only include the code below to make a few obs
to the slides. please do not use head() on your assignments
-->

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
using echo: false here is just to make the slides cleaner.
-->

```{r}
#| echo: false
clubhouse %>% 
  mutate(goofy_laughs = median(goofy_laughs),
         donald_grumbles = median(donald_grumbles)) %>% 
  select(time_with_friends, goofy_laughs, donald_grumbles) %>%
  head()
```    

## Plotting using `library(ggplot2)`

- Then, constructing our graph,
    
```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_k3), color = "#004C97", size = 1.5) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Plotting using `library(ggplot2)`
    
```{r}
#| echo: false
#| fig-align: 'center'
(k3 <- clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_k3), color = "#004C97", size = 1.5) +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw())
```

## Let's Explore...

- Hm... what if we put the three lines on top of one another? How different are the adjusted slopes?


```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = time_with_friends,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3, alpha = .15) + 
  geom_line(aes(y = predicted_happiness_k1), color = "black", size = 1) +
  geom_line(aes(y = predicted_happiness_k2), color = "blue", size = 1) +
  geom_line(aes(y = predicted_happiness_k3), color = "red", size = 1) +
  annotate("text", x = .5, y = 38, label = "k=1", color = "black") +
  annotate("text", x = .5, y = 35, label = "k=2", color = "blue") +
  annotate("text", x = .5, y = 32, label = "k=3", color = "red") +
  labs(x = "Time Spent with Friends (minutes)",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Example 4: Multiple Regression (*k* = 3)

- For our fourth example, let's again return to our full model.

- We looked at clubhouse happiness (*clubhouse_happiness*) as a function of time spent with friends (*time_with_friends*), big, goofy laughs (*goofy_laughs*), and how much Donald grumbles (*donald_grumbles*).

```{r}
m4 <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles,
          family = "gaussian",
          data = clubhouse)
m4 %>% tidy()
```

$$
\hat{\text{happiness}} = 47.58 + 3.58 \text{ time} + 0.66 \text{ laughs} - 1.06 \text{ grumbles}
$$

## Model Visualization: Multiple Regression

- Let's now consider the relationship between clubhouse happiness and Donald's grumbles.

    - Donald grumbles will be on the $x$-axis and allowed to vary.
    - We will hold goofy laughs constant at some value.
    - We will also hold time spent with friends constant at some value.
    
## Model Visualization: Multiple Regression

```{r}
clubhouse <- clubhouse %>%
  mutate(predicted_happiness_d = 47.58 + 3.58*median(time_with_friends) + 0.66*median(goofy_laughs) - 1.06*donald_grumbles)
```

<!-- 
note: I only include the code below to make a few obs
to the slides. please do not use head() on your assignments
-->

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
using echo: false here is just to make the slides cleaner.
-->

```{r}
#| echo: false
clubhouse %>% 
  mutate(goofy_laughs = median(goofy_laughs),
         time_with_friends = median(time_with_friends)) %>% 
  select(time_with_friends, goofy_laughs, donald_grumbles) %>%
  head()
```       

## Plotting using `library(ggplot2)`

- Then, constructing our graph,
    
```{r}
#| eval: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = donald_grumbles,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_d), color = "#004C97", size = 1.5) +
  labs(x = "Number of Donald Grumbles",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Plotting using `library(ggplot2)`

- Then, constructing our graph,
    
```{r}
#| echo: false
#| fig-align: 'center'
clubhouse %>% ggplot(aes(x = donald_grumbles,
                         y = clubhouse_happiness)) +
  geom_point(color = "#009CDE", size = 3) + 
  geom_line(aes(y = predicted_happiness_d), color = "#004C97", size = 1.5) +
  labs(x = "Number of Donald Grumbles",
       y = "Clubhouse Happiness") +
  theme_bw()
```

## Wrap Up

- In this lecture, we explored how to visualize simple and multiple linear regression models using the `ggplot2` library.

- For simple linear regression, we visualized the relationship between the outcome and predictor using a scatterplot and regression line.

- For multiple linear regression, we visualized the relationship between the outcome and one predictor while holding the other predictors constant.

- Every week, we will review model visualization.

    - The general ideas won't change, but things will get tricky when we add categorical predictors and leave the normal distribution.
    
- Next lecture: Model Assumptions and Diagnostics    
