---
title: "Review of Linear Regression"
execute:
  echo: true
  warning: false
  message: false
format: 
  revealjs:
    theme: default
    self-contained: true
    slide-number: true
    width: 1600
    height: 900
    df-print: paged
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

## Introduction

- Recall the general linear model, $$y = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k$$

- This is a multiple regression model because it has multiple predictors ($x_i$).

    - A special case is simple linear regression, when there is a single predictor. $$y = \beta_0 + \beta_1 x_1 $$

- $\beta_0$ is the $y$-intercept, or the average outcome ($y$) when all $x_i = 0$.

- $\beta_i$ is the slope for predictor $i$ and describes the relationship between the predictor and the outcome, after adjusting (or accounting) for the other predictors in the model.

## Lecture Example Set Up

- On a busy day at the clubhouse, Mickey Mouse wants to understand what drives "happiness" at the end of the day. For each day, he records (in the <u>**clubhouse**</u> dataset):

    - Time with friends (in hours; *time_with_friends*): how many hours Mickey spends hanging out with his friends.
    - Goofy Laughs (a count; *goofy_laughs*) – how many big goofy laughs happen that day.
    - Donald Grumbles (a count; *donald_grumbles*): how many times Donald gets frustrated and grumbles.
    - Clubhouse Happiness (a score; *clubhouse_happiness*): an overall happiness score at the end of the day.

- In this lecture, we will use a linear model to explore the relationships between happiness and laughs, grumbles, and time spent with friends.

## Lecture Example Set Up 

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
This step is not necessary on your assignments, just for clarity on slides    
-->

```{r}
#| echo: false
library(tidyverse)
library(broom)
```

- Let's pull in our data, 

<!-- read_csv() is from library(tidyverse) -->
<!-- head() is to display data on the slide -->

```{r}
clubhouse <- read_csv("~/Desktop/GitHub/samanthaseals/STA4231_STA6262/files/data/lectures/W1_mickey_clubhouse.csv")
head(clubhouse)
```

## Constructing the Linear Model (R)

- We will use the `glm()` function to construct our linear model.

    - `glm()` stands for Generalized Linear Model.
    - For linear regression, we will specify `family = gaussian`.

<!--
If you are looking at the raw .qmd, note that
the following code chunk has the option "eval: false" so that
the code is displayed but is not evaluated.
This is entirely for slide purposes; you will not use
#| eval: false in your assignments.
-->

```{r}
#| eval: false
m <- glm(y ~ x_1 + x_2 + ... + x_k,
         family = "gaussian",
         data = dataset_name)
```

## Constructing the Linear Model 

- In our example,

```{r}
m <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles,
         family = gaussian,
         data = clubhouse)
tidy(m)
```

$$
\begin{align*}
\hat{y} &= 47.58 + 3.58 \ x_1 + 0.65 \ x_2 - 1.06 \ x_3 \\
\hat{\text{happiness}} &= 47.58 + 3.58 \text{ time} + 0.65 \text{ laughs} - 1.06 \text{ grumbles}
\end{align*}
$$

## Interpretation of Slope

- We want to put the slope into perspective for whoever we are collaborating with.

- Basic interpretation: for every 1 [units of $x_i$] increase in [$x_i$], [$y$] [increases or decreases] by $\left[ \left| \hat{\beta}_i \right| \right]$ [units of $y$].

    - We say that $y$ is decreasing if $\hat{\beta}_i < 0$ and $y$ is increasing if $\hat{\beta}_i > 0$.
    
- Note that in the case of multiple regression, there is an unspoken "after adjusting for everything else in the model" at the end of the sentence.

    - Always remember that we are looking at the relationship between $y$ and $x_i$ after adjusting for all other predictors included in the predictor set.    

## Interpretation of Slope

- Let's interpret the slopes for the model we constructed.

```{r}
m %>% tidy()
```

## Interpretation of Slope

- Let's interpret the slopes for the model we constructed.

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
This step is not necessary on your assignments, just for clarity on slides    
-->

<!-- 
Please also note that I am using the coefficients()
function so that I can display the slopes without the full table.
This is not necessary on your assignments.
-->

```{r}
#| echo: false
coefficients(m)[-1]
```

- For every 1 hour increase in time spent with friends, clubhouse happiness increases by 3.58 points, after adjusting for goofy laughs and Donald grumbles.

- For every 1 big goofy laugh increase, clubhouse happiness increases by 0.65 points, after adjusting for time spent with friends and Donald grumbles.

- For every 1 Donald grumble increase, clubhouse happiness decreases by 1.06 points, after adjusting for time spent with friends and goofy laughs.

## Interpretation of Slope

- We can also scale our interpretations. e.g.,

    - For every k [units of $x_i$] increase in [$x_i$], [$y$] [increases or decreases] by $\left[ k \times \left| \hat{\beta}_i \right| \right]$ [units of $y$].
    
- In our example,

    - For every increase of 1 Donald grumble, clubhouse happiness decreases by 1.06 points, after adjusting for time spent with friends and goofy laughs.
    - For every increase of 5 Donald grumbles, clubhouse happiness decreases by 5.3 points, after adjusting for time spent with friends and goofy laughs.

## Interpretation of Intercept

- The intercept is the average of the outcome when all predictors in the model are equal to 0. 

    - We can think of this as the "baseline" level of the outcome before any predictors have an effect.

- In our example,

<!-- 
If you are looking at the raw .qmd, note that
the following code chunk has the option "echo: false" so that
the output is displayed but the code itself is hidden. 
This step is not necessary on your assignments, just for clarity on slides    
-->

<!-- 
Please also note that I am using the coefficients()
function so that I can display the slopes without the full table.
This is not necessary on your assignments.
-->

```{r}
#| echo: false
coefficients(m)[1]
```

- The average clubhouse happiness when time with friends = 0 hours, goofy laughs = 0, and Donald grumbles = 0 is 47.58 points.

## Confidence Intervals for $\beta_i$ (R)

- Recall confidence intervals -- they allow us to determine how "good" our estimation is.

- In general, CIs will take the form
<center> point estimate $\pm$ margin of error </center>
  
- The margin of error is a critical value (e.g., $t_{1-\alpha/2}$) multiplied by the standard error of the point estimate.

- Recall that the standard error accounts for the sample size.

- In R, we will run the model results through the `tidy()` function, but ask for the confidence intervals. 

<!--
If you are looking at the raw .qmd, note that
the following code chunk has the option "eval: false" so that
the code is displayed but is not evaluated.
This is entirely for slide purposes; you will not use
#| eval: false in your assignments.
-->

```{r}
#| eval: false
m %>% tidy(conf.int = TRUE)
```

## Confidence Intervals for $\beta_i$

- In our example, 

```{r}
m %>% tidy(conf.int = TRUE)
```

- We have the following CIs:
  
    - 95% CI for $\beta_{\text{time}}$ is (2.37, 4.78)
    - 95% CI for $\beta_{\text{goofy}}$ is (0.52, 0.79)
    - 95% CI for $\beta_{\text{Donald}}$ is (-1.34, -0.79)
    
## Confidence Intervals for $\beta_i$

- What if we want something other than a 95% CI?

<!--
If you are looking at the raw .qmd, note that
the following code chunk has the option "eval: false" so that
the code is displayed but is not evaluated.
This is entirely for slide purposes; you will not use
#| eval: false in your assignments.
-->

```{r}
#| eval: false
m %>% tidy(conf.int = TRUE, conf.level = insert_level_here)
```

## Confidence Intervals for $\beta_i$

- In our example,

```{r}
m %>% tidy(conf.int = TRUE, conf.level = 0.99)
```
- We have the following CIs:
  
    - 99% CI for $\beta_{\text{time}}$ is (1.99, 5.16)
    - 99% CI for $\beta_{\text{goofy}}$ is (0.48, 0.84)
    - 99% CI for $\beta_{\text{Donald}}$ is (-1.42, -0.71)    
    
## Confidence Intervals for $\beta_i$

- Let's compare the two sets of CIs.

- 95% CIs for $\beta_i$:
  
    - 95% CI for $\beta_{\text{time}}$ is (2.37, 4.78)
    - 95% CI for $\beta_{\text{goofy}}$ is (0.52, 0.79)
    - 95% CI for $\beta_{\text{Donald}}$ is (-1.34, -0.79)
    
- 99% CIs for $\beta_i$:
  
    - 99% CI for $\beta_{\text{time}}$ is (1.99, 5.16)
    - 99% CI for $\beta_{\text{goofy}}$ is (0.48, 0.84)
    - 99% CI for $\beta_{\text{Donald}}$ is (-1.42, -0.71)        

## Significant Regression Line

- We now will ask ourselves if any of our slopes are flat.

$$
H_0: \ \beta_1 = \beta_2 = ... = \beta_k = 0
$$

- What does it mean if the slopes are flat?

    - If a slope is flat, it means that there is no relationship between that predictor and the outcome, after adjusting for the other predictors in the model.
        
- We will use hypothesis testing to determine if at least one slope is non-zero.

    - This is often referred to as the omnibus $F$-test in linear regression.
    
    - We will instead use the likelihood ratio test.
        
        - This is because we will continue to use this test outside of the normal distribution.

## Significant Regression Line

- For any model constructed, we can compute deviance.

    - Deviance measures the unexplained variability in the outcome under the fitted model.
    - Lower deviance means the model fits the data better.
        - The actual value is not meaningful.
    
- To determine the test statistic, we examine the difference between two models:

    - Full model: includes all predictors.
    - Reduced model: includes only the intercept (no predictors).    

- Then, we look at the difference between the deviances of the full and reduced models.
	
    - This difference has an approximate $\chi^2$ distribution.

## Significant Regression Line

- **Hypotheses**
  
    - $H_0: \ \beta_1 = ... = \beta_k = 0$ 
    - $H_1:$ at least one $\beta_i \ne 0$ 
  
- **Test Statistic and *p*-Value**
  
    - $\chi^2_0 = [\text{value from R}]$, $p = [\text{value from R}]$

- **Rejection Region**
  
    - Reject $H_0$ if $p < \alpha$; $\alpha = [\text{assumed } \alpha]$.
    
- **Conclusion/Interpretation**

    - [Reject (*if $p < \alpha$*) or fail to reject (*if $p \ge \alpha$*)] $H_0$. There [is (*if reject*) or is not (*if FTR*)] sufficient evidence to suggest that at least one slope is non-zero.

<!-- i.e.,
Reject the null. There is sufficient evdience to suggest...
Fail to reject the null. There is not sufficient evdience to suggest...
-->
    
## Significant Regression Line (R)

- We will *compare models* to determine the significance of the line.

    - Full model: includes all predictors.
    - Reduced model: includes only the intercept (no predictors). 
    
- In R,    

```{r}
#| eval: false
full <- glm(y ~ x_1 + x_2 + ... + x_k, data = dataset_name, family = "gaussian")
reduced <- glm(y ~ 1, data = dataset_name, family = "gaussian")
anova(reduced, full, test = "LRT")
```        

## Significant Regression Line 

- In our example,    

```{r}
full <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles,
            data = clubhouse,
            family = "gaussian")
reduced <- glm(clubhouse_happiness ~ 1, 
               data = clubhouse, 
               family = "gaussian")
anova(reduced, full, test = "LRT")
```  

## Significant Regression Line

- **Hypotheses**
  
    - $H_0: \ \beta_{\text{time}} = \beta_{\text{goofy}} = \beta_{\text{Donald}} = 0$ 
    - $H_1:$ at least one $\beta_i \ne 0$ 
  
- **Test Statistic and *p*-Value**
  
    - $\chi^2_0 = 17305$, $p < 0.001$

- **Rejection Region**
  
    - Reject $H_0$ if $p < \alpha$; $\alpha = 0.05$.
    
- **Conclusion/Interpretation**

    - Reject $H_0$. There is sufficient evidence to suggest that at least one slope is non-zero.
    
## Significant Predictors of $y$

- **Hypotheses**

    - $H_0: \ \beta_i  = 0$ 
    - $H_1: \ \beta_i \ne 0$ 

- **Test Statistic and *p*-Value**

    - $t_0 = [\text{value from R}]$, $p = [\text{value from R}]$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; ; $\alpha = [\text{assumed } \alpha]$.
    
- **Conclusion/Interpretation**

    - [Reject (*if $p < \alpha$*) or fail to reject (*if $p \ge \alpha$*)] $H_0$. There [is (*if reject*) or is not (*if FTR*)] sufficient evidence to suggest that [*predictor name*] significantly predicts [*outcome*], after adjusting for the other predictors in the model.
    
## Significant Predictors of $y$ (R)

- Because we currently are only dealing with continuous predictors, we will use the results from `tidy()`.

<!--
If you are looking at the raw .qmd, note that
the following code chunk has the option "eval: false" so that
the code is displayed but is not evaluated.
This is entirely for slide purposes; you will not use
#| eval: false in your assignments.
-->

```{r}
#| eval: false
m %>% tidy()
```    

## Significant Predictors of $y$

- In our example,

```{r}
m %>% tidy()
```   

## Significant Predictors of $y$

- **Hypotheses**

    - $H_0: \ \beta_{\text{time}}  = 0$ 
    - $H_1: \ \beta_{\text{time}} \ne 0$ 
- **Test Statistic and *p*-Value**

    - $t_0 = 5.80$, $p < 0.001$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; ; $\alpha = 0.05$.   
    
- **Conclusion/Interpretation**

    - Reject $H_0$. There is sufficient evidence to suggest that time with friends significantly predicts the clubhouse happiness, after adjusting for the other predictors in the model.
    
## Significant Predictors of $y$

- **Hypotheses**

    - $H_0: \ \beta_{\text{goofy}}  = 0$ 
    - $H_1: \ \beta_{\text{goofy}} \ne 0$ 

- **Test Statistic and *p*-Value**

    - $t_0 = 9.55$, $p < 0.001$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; ; $\alpha = 0.05$.  
    
- **Conclusion/Interpretation**

    - Reject $H_0$. There is sufficient evidence to suggest that goofy laughs significantly predict the clubhouse happiness, after adjusting for the other predictors in the model.
    
    
## Significant Predictors of $y$

- **Hypotheses**

    - $H_0: \ \beta_{\text{Donald}}  = 0$ 
    - $H_1: \ \beta_{\text{Donald}} \ne 0$ 

- **Test Statistic and *p*-Value**

    - $t_0 = -7.66$, $p < 0.001$

- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; ; $\alpha = 0.05$.    
    
- **Conclusion/Interpretation**

    - Reject $H_0$. There is sufficient evidence to suggest that Donald's grumbles significantly predict the clubhouse happiness, after adjusting for the other predictors in the model.

## Reporting Results


- I typically use a table to provide results to collaborators.

| Predictor | $\hat{\beta}_i$ (95% CI for $\beta_i$) | *p*-value |
|:--|:--:|--:|
| time with friends | 3.58 (2.37, 4.78) | < 0.001 |
| goofy laughs | 0.65 (0.52, 0.80) | < 0.001 |
| Donald grumbles | -1.06 (-1.34, -0.79) | < 0.001 |

- As time with friends increases by 1 minute, clubhouse happiness increases by 3.58 points.
- As goofy laughs increase by 1, clubhouse happiness increases by 0.65 points.
- As Donald grumbles increase by 1, clubhouse happiness decreases by 1.06 points

## Wrap Up

- This lecture reviewed the key components of linear regression:

    - Constructing the model in R
    - Interpreting slopes and intercepts
    - Constructing confidence intervals for slopes
    - Testing for a significant regression line
    - Testing for significant predictors
    
- Next lecture: Model Visualization