---
title: "Interaction Terms: <br> Continuous x Continuous"
execute:
  echo: true
  warning: false
  message: false
format: 
  revealjs:
    theme: default
    self-contained: true
    slide-number: true
    width: 1600
    height: 900
    df-print: paged
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

## Introduction 

- Recall the general linear model, 

$$ y = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k + \varepsilon $$

- Until now, we have talked about models with only main effects.

    - e.g., $x_1$, $x_2$

- Today, we will begin talking about interactions.

    - e.g., $x_1 \times x_2$

## Interactions

- Recall interactions from two-way ANOVA:

    - The relationship between the outcome and one predictor depends on the level of another predictor.

- Interactions work (and are specified) the same way in regression. 
- The usual caveats apply:

    - We do not want to load models with too many interactions.
    
    - We favor simplicity over interactions that do not add much to the predictive power of the model.

    - We do not want higher than two-way interactions unless necessary.
    
- In this lecture, we will focus on continuous $\times$ continuous interactions.

## Interactions 

- We will construct what is called a hierarchical well-formulated (HWF) model. 

- This means that when a higher-order interaction term is included in the model, all lower-order terms are also included.

    - e.g., when a two-way interaction is included, we also include the corresponding main effects. $$y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_1 x_2$$

    - e.g., when a three-way interaction is included, we also include the corresponding main effects and two-way interactions. $$y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3 + \beta_4 x_1 x_2 + \beta_5 x_1 x_3 + \beta_6 x_2 x_3 + \beta_7 x_1 x_2 x_3$$

## Lecture Example Set Up

```{r}
#| echo: false
library(tidyverse)
library(ssstats)
library(broom)
library(kableExtra)
```

- On a busy day at the clubhouse, Mickey Mouse wants to understand what drives "happiness" at the end of the day. For each day, he records (in the <u>**clubhouse**</u> dataset):

    - Time with friends (in hours; *time_with_friends*): how many hours Mickey spends hanging out with his friends.
    - Goofy Laughs (a count; *goofy_laughs*) – how many big goofy laughs happen that day.
    - Donald Grumbles (a count; *donald_grumbles*): how many times Donald gets frustrated and grumbles.
    - Clubhouse Happiness (a score; *clubhouse_happiness*): an overall happiness score at the end of the day.
    
```{r}
clubhouse <- read_csv("https://raw.githubusercontent.com/samanthaseals/SDSII/refs/heads/main/files/data/lectures/W1_mickey_clubhouse.csv")
```

## Example 1

- Let's consider a basic example,

```{r}
m1 <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + time_with_friends:goofy_laughs,
         family = "gaussian",
         data = clubhouse)
m1 %>% tidy()
```

$$
\hat{\text{happiness}} = 20.13 + 7.78 \text{ time} + 1.30 \text{ laughs} - 0.16 \ (\text{time$\times$laughs})
$$

## Interpretations

- How do we interpret the interaction term?

- This is the change in the effect of one predictor for a one-unit increase in the other predictor.

    - You may also see this referred to as a "modifier" -- it is *modifying* the slope for $x_1$ based on the value of $x_2$ (and vice versa).
    
- In an example without context, suppose we have the model,

$$
\hat{y} = 5.2 + 2.3 x_1 + 4.5 x_2 - 1.2 (x_1 \times x_2)
$$

- The coefficient for the interaction term (-1.2) indicates that for every one unit increase in $x_2$, the effect of $x_1$ on $y$ decreases by 1.2 units.

    - Similarly, for every one-unit increase in $x_1$, the effect of $x_2$ on $y$ decreases by 1.2 units.

## Example 1

- Consider our clubhouse example again,

$$
\hat{\text{happiness}} = 20.13 + 7.78 \text{ time} + 1.30 \text{ laughs} - 0.16 \ (\text{time$\times$laughs})
$$

- The interaction term tells us that:

    - For every additional goofy laugh, the effect of time with friends on clubhouse happiness decreases by 0.16 units.
    - For every additional hour spent with friends, the effect of goofy laughs on clubhouse happiness decreases by 0.16 units.
    
## Testing the Interaction Term

- In the case of a continuous $\times$ continuous interaction, we can use the individual *p*-value (from `tidy()`) to determine statistical significance.

    - This is the *same test* we would use for any other individual coefficient in the model.

- **Hypotheses**

    - $H_0$: $\beta_{x_i \times x_j} = 0$
    - $H_1$: $\beta_{x_i \times x_j} \ne 0$

- **Test Statistic & *p*-Value**

    - $t_0 =$ (from `tidy()`)
    - $p$ = (from `tidy()`)

## Example 1

- Looking at this in our example,

```{r}
m1 %>% tidy()
```

- The *p*-value for the interaction term is 0.070.

- The interaction is *not significant* and should not be included in our model. 

## Example 1

- **Hypotheses**

    - $H_0$: $\beta_{3} = 0$
    - $H_1$: $\beta_{3} \ne 0$

- **Test Statistic & *p*-Value**

    - $t_0 = - 1.819$
    - $p = 0.070$  
    
- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha = 0.05$
    
- **Conclusion & Interpretation**

    - FTR $H_0$. The slope of time with friends on clubhouse happiness does not significantly depend on the number of goofy laughs (and vice versa).

## Syntax when Specifying Interactions

- There are two ways we can specify interactions in our model:

    - Using the `:` operator to specify each interaction term individually.
    - Using the `*` operator to specify all main effects and interactions at once.
    
- Consider a model with time with friends ($x_1$), goofy laughs ($x_2$), Donald grumbles ($x_3$), and all possible interactions ($x_1 x_2$, $x_1 x_3$, $x_2 x_3$, $x_1 x_2 x_3$).    
    
    - When we use the `:` operator, <center>$y \sim x_1 + x_2 + x_3 + x_1:x_2 + x_1:x_3 + x_2:x_3 + x_1:x_2:x_3$</center>

    - When we use the `*` operator, <center>$y \sim x_1 * x_2 * x_3$</center>

- Obviously this is faster, but remember that it gives us less control. 

    - e.g., if we decide we do not want the three-way interaction, we will have to completely respecify our model in the code vs. just deleting the term we want to remove.
    
## Example 2

- Applying this to our data,

- Using the `:` operator,

```{r}
m2a <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles +
             time_with_friends:goofy_laughs + time_with_friends:donald_grumbles + goofy_laughs:donald_grumbles +
             time_with_friends:goofy_laughs:donald_grumbles,
         family = "gaussian",
         data = clubhouse)
```

- Using the `*` operator,
```{r}
m2b <- glm(clubhouse_happiness ~ time_with_friends * goofy_laughs * donald_grumbles,
         family = "gaussian",
         data = clubhouse)
```

- If indeed the model specifications are equivalent, we should see the same results from `tidy()`.

## Example 2

- Looking at the results,

```{r}
m2a %>% tidy()
```

## Example 2

- Looking at the results,

```{r}
m2b %>% tidy()
```

## Example 2

- Some fancy code is in the .qmd file to show you how this table was constructed, but here are the results side-by-side:

```{r}
#| echo: false

m2a_results <- tidy(m2a, conf.int = TRUE)
m2b_results <- tidy(m2b, conf.int = TRUE)
library(dplyr)
library(tidyr)
m2a_results <- m2a_results %>%
  select(term, estimate, conf.low, conf.high, p.value) %>%
  rename(m2a_estimate = estimate,
         m2a_conf_low = conf.low,
         m2a_conf_high = conf.high)
m2b_results <- m2b_results %>%
  select(term, estimate, conf.low, conf.high, p.value) %>%
  rename(m2b_estimate = estimate,
         m2b_conf_low = conf.low,
         m2b_conf_high = conf.high)
combined <- m2a_results %>%
  full_join(m2b_results, by = "term") %>%
  mutate(`: operator` = paste0(round(m2a_estimate, 2), " (",
                              round(m2a_conf_low, 2), ", ",
                              round(m2a_conf_high, 2), ")"),
         `* operator` = paste0(round(m2b_estimate, 2), " (",
                              round(m2b_conf_low, 2), ", ",
                              round(m2b_conf_high, 2), ")")) %>%
  select(term, `: operator`, `* operator`)
kable(combined)
```

## When to Remove Interaction Terms

- When do I remove interaction terms from my models?

    - If the interaction term is not statistically significant, you can consider removing it from the model. Removing any non-significant interactions will make the model easier to interpret.
    
- *However*, be cautious when removing interaction terms. This may require discussion with the team.

    - If there is a theoretical reason to keep the interaction term, the team may choose to keep it in the model even if it is not statistically significant.
    
- Because I always want the option of removing terms easily, that is why I prefer to use the `:` operator when specifying interactions.

## Example 2

- Returning to our example with a three-way interaction,

```{r}
m2a %>% tidy()
```

## Example 2

- **Hypotheses**

    - $H_0$: $\beta_{7} = 0$
    - $H_1$: $\beta_{7} \ne 0$

- **Test Statistic & *p*-Value**

    - $t_0 = - 0.983$
    - $p = 0.326$  
    
- **Rejection Region**

    - Reject $H_0$ if $p < \alpha$; $\alpha = 0.05$
    
- **Conclusion & Interpretation**

    - FTR $H_0$. The three-way interaction term is not statistically significant, meaning that the two-way interactions *do not depend* on the level of the third predictor.
    
## Example 2

- Removing the three-way interaction term,

```{r}
m2c <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles + 
             time_with_friends:goofy_laughs + time_with_friends:donald_grumbles + goofy_laughs:donald_grumbles,
         family = "gaussian",
         data = clubhouse)
m2c %>% tidy()
```

## Testing Multiple Interactions at Once 

- When I have multiple (two-way) interactions in my model, I will sometimes perform a partial $F$ test to assess the significance of all interaction terms at once.

    - This is the full vs. reduced model approach we learned in multiple regression for the purpose of testing the overall significance of the model.
    
- The corresponding hypotheses are:
    
    - $H_0$: $\beta_{\text{int}_1} = \beta_{\text{int}_2} = ... = \beta_{\text{int}_m} = 0$ (all interaction terms are zero)
    - $H_1$: At least one $\beta_{\text{int}_k} \ne 0$ (at least one interaction term is non-zero)
    
- Then, in R, we take a similar approach to what we saw before,    
    
```{r}
#| eval: false
full <- glm(y ~ x_1 + x_2 + x_3 + x_1:x_2 + x_1:x_3 + x_2:x_3,
         family = "gaussian",
         data = data)
reduced <- glm(y ~ x_1 + x_2 + x_3,
           family = "gaussian",
           data = data)
anova(reduced, full, test = "F")
```

## Example 2

- Returning to our example, let's test all three two-way interactions at once.

```{r}
full <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles + 
             time_with_friends:goofy_laughs + time_with_friends:donald_grumbles + goofy_laughs:donald_grumbles,
         family = "gaussian",
         data = clubhouse)
reduced <- glm(clubhouse_happiness ~ time_with_friends + goofy_laughs + donald_grumbles,
         family = "gaussian",
         data = clubhouse)
anova(reduced, full, test = "F")
```

## Example 2

- We fail to reject, meaning we can assume all of the interaction-related slopes are 0.

```{r}
anova(reduced, full, test = "F")
```

- This gives me *justification* in removing all of the two-way interactions *at the same time*.

## "Backward Selection"

- True backward selection would involve:

    - Starting with a full model with all possible interaction terms.
    - Testing the highest-order interaction term first.
    - If non-significant, remove it and refit the model.
    - Repeat until all remaining interaction terms are significant.
    - If no significant interactions, continue to test and remove main effects as needed.
    
## "Backward Selection"
    
- My version of backward selection:

    - Start with a full model with interaction terms *of interest*.
    - Test all interaction terms at once (partial $F$ test).
    - If non-significant, remove all interaction terms and refit the model.
    - If significant, test each interaction term individually and remove non-significant ones.
    - Repeat until all remaining interaction terms are significant OR all interaction terms have been removed.
    
## "Backward Selection"

- This is not a perfect method, but it is a practical way to simplify models with interaction terms.

    - I am in favor of *parsimony*. I will always recommend a simpler model over a more complex one when there is "not a difference". 

- Note that because we are working with hierarchical well-formulated models, if main effects are involved in interactions, they must remain in the model as main effects.

    - y ~ x1 + x2 + x3 + x1:x2 + x1:x3 $\to$ no main effects can be removed.
    
    - y ~ x1 + x2 + x1:x2 + x1:x3 $\to$ not a valid model!
    
- If a main effect is *not* involved in an interaction, it can be removed *if desired*.

    - y ~ x1 + x2 + x3 + x1:x2 + $\to$ $x_3$ can be removed if desired.
    
## Digging Deeper into Interactions

- Let's solidify our understanding of interaction terms.

- Let's return to Example 1, 

$$
\hat{\text{happiness}} = 20.13 + 7.78 \text{ time} + 1.30 \text{ laughs} - 0.16 \ (\text{time$\times$laughs})
$$

- To better understand the interaction, we can compute *simple slopes*.

    - These are the slopes of one predictor at specific values of the other predictor.
    
- What is the slope for time with friends when goofy laughs = 0? 25? 50?


## Example 1

- Starting model,

$$
\hat{\text{happiness}} = 20.13 + 7.78 \text{ time} + 1.30 \text{ laughs} - 0.16 \ (\text{time$\times$laughs})
$$

- Plugging in goofy laughs = 0,

$$
\begin{align*}
\hat{\text{happiness}} &= 20.13 + 7.78 \text{ time} + 1.30(0) - 0.16 \ (\text{time$\times$0}) \\
&= 20.13 + 7.78 \text{ time} 
\end{align*}
$$

- When there are no goofy laughs, the slope for time with friends is 7.78. This means that as time increases by one hour, happiness increases by 7.78 points.

## Example 1

- Starting model,

$$
\hat{\text{happiness}} = 20.13 + 7.78 \text{ time} + 1.30 \text{ laughs} - 0.16 \ (\text{time$\times$laughs})
$$

- Plugging in goofy laughs = 25,

$$
\begin{align*}
\hat{\text{happiness}} &= 20.13 + 7.78 \text{ time} + 1.30(25) - 0.16 \ (\text{time$\times$25}) \\
&= 52.63 + 3.78 \text{ time} 
\end{align*}
$$

- When there are 25 goofy laughs, the slope for time with friends is 3.78. This means that as time increases by one hour, happiness increases by 3.78 points.

## Digging Deeper into Interactions

- Let's look at these together,

    - When there are no goofy laughs, the slope for time with friends is 7.78. This means that as time increases by one hour, happiness increases by 7.78 points.
    
    - When there are 25 goofy laughs, the slope for time with friends is 3.78. This means that as time increases by one hour, happiness increases by 3.78 points.
    
- Can we make an overall statement?

    - As the number of goofy laughs increases, the effect of time with friends on happiness decreases. This is consistent with the negative interaction term (-0.16). 
    
    - How would I explain it to Mickey? When the number of laughs increases, it <u>takes longer</u> to increase the clubhouse happiness level. 
    
## Digging Deeper into Interactions

- Let's return to the initial slide for this section.

$$
\hat{\text{happiness}} = 20.13 + 7.78 \text{ time} + 1.30 \text{ laughs} - 0.16 \ (\text{time$\times$laughs})
$$

- To better understand the interaction, we can compute *simple slopes*.

    - These are the slopes of one predictor at specific values of the other predictor.
    
- What is the slope for time with friends when goofy laughs = 0? 25? 50?

- Note!

    - We could easily have asked, what is the slope for goofy laughs when time with friends = 0? 2? 4? 
    
## Digging Deeper into Interactions

- How do I choose what to plug in for?

    - I look at the research question / narrative that we have to this point. Sometimes it is obvious which one the "story" is about. 
    
    - The "interesting variable" is the one that we will allow to be modified by the interaction.
    
- What if I can't tell or both are of interest?

    - I do not decide. I will ask the research team for clarification.
    
    - Sometimes I will compute one simple slope each way to give a concrete example.
    
- Always remember to underscore that this is for *interpreting* the model -- it does not change the model itself!    

## Wrap Up

- This lecture introduced the concepts of interaction terms in our models.

- We started with the continuous $\times$ continuous interaction type to get comfortable with the idea of interactions in regression.

- As we can see, relaying the meaning of interaction terms can be tricky and, depending on number and type of interaction term, can very steeply increase the complexity of the model.

- In the next lecture, we will discuss categorical $\times$ categorical interactions.

    - We will see where things are both easier and trickier than continuous $\times$ continuous interactions, unfortunately.

