---
output: 
  beamer_presentation:
    fig_caption: false
    includes:
      in_header: header.tex
classoption: 
  - "aspectratio=169"
  
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../lectures/PDF") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.path = "images")

library(tidyverse)
library(fastDummies)
library(gsheet)
library(car)
library(pscl)

data <- as_tibble(gsheet2tbl("https://docs.google.com/spreadsheets/d/1JoI0AffzZdGkW3FCCWHgRaZeuGrBt1Qa-WMT3_l8XCA/edit?usp=sharing"))
```


## {.standout} 
\vskip12em
\begin{center}{\color{white} \huge \textbf{Zero-Inflated Models}} \vskip1em
{\color{white} \Large Statistics for Data Science II}
\end{center}

## Introduction

\vskip1em
Now that we know how to handle count data, we need to discuss what to do if there are "too many" zeros in the response.
\begin{itemize}
  \item If we look at a histogram or bar chart of the responses, we may see a ``spike'' at zero -- this spike indicates zero inflation.
\end{itemize}

The zero-inflated Poisson model is as follows:
\[  y_i \sim \left\{
\begin{array}{ll}
      0, & \text{with probability }1 - \phi_i \\
      \text{Poisson}(\lambda_i), & \text{with probability } \phi_i \\
\end{array} 
\right. \]

Thus, we are jointly modeling (1) a logit model for predicting excess zeros, (2) the Poisson count model.

## Modeling

\vskip1em
We will specify this in R using the \texttt{zeroinfl()} function. \vskip.5em

e.g., \texttt{zeroinfl(outcome $\sim$ predCount1 + predCount2 + ... | predZero1 + predZero2 + ..., data = dataset)} 

\begin{itemize}
  \item Note that we do not need the same predictors in each model.
\end{itemize}

\textbf{Example:} \vskip1em

```{r}
m1 <- zeroinfl(satellites_num ~ width_cm + spine_cond +
            width_cm:spine_cond | width_cm + spine_cond,
          data=data)
```

## Modeling
\vskip1em
\textbf{Example:} \vskip.5em
\scriptsize
```{r}
summary(m1)[1]
```
## Modeling
 
\vskip1em
\textbf{Example}:
\begin{itemize}
  \item The resulting models are
\end{itemize} \vskip-2em
\begin{align*}
\ln \left( \hat{Y}_i \right) &= 2.94 - 0.04 \text{width} - 1.01 \text{spine} + 0.03 (\text{width $\times$ spine}) \\
\text{ln}\left( \frac{\hat{\pi}_i}{1-\hat{\pi}_i}  \right) &= 12.50 - 0.50 \text{width} - 0.04 \text{spine}
\end{align*}
\begin{itemize}
  \item $Y_i$ is the number of satellites, 
  \item $\pi_i = P[Z = 1]$, where
  \begin{itemize}
    \item \vskip-2em \begin{flalign*}
  Z_i &= \left\{
  \begin{array}{ll}
        0 & \text{ if } Y_i = 0 \\
        1 & \text{ if } Y_i > 0 
  \end{array} 
  \right. &
  \end{flalign*}
\end{itemize}
\end{itemize}

## Interpretations

Because we are really interested in modeling the count data ($Y$), we will focus on interpreting the Poisson regression. 

Like before, we will convert the $\hat{\beta}_i$ to IRR$_i$ and interpret in terms of the multiplicative effect.

See previous lectures for interpretations in models for count data.

## Inference

\vskip2em
Testing for significant predictors is the same as previously discussed. 

\textbf{Example:}
\scriptsize
```{r}
summary(m1)[1]
```

## Inference

\vskip1em
Constructing confidence intervals is the same as previously discussed. \vskip.5em

\textbf{Example:}
\scriptsize
```{r}
confint(m1)
```

## Modeling - Negative Binomial

Recall that Poisson regression is not always appropriate. When this is the case, we will use the zero-inflated negative binomial.

We are still using the \texttt{zeroinfl()} function, but now we specify \texttt{dist = "negbin"}.

\textbf{Example:}

```{r}
m2 <- zeroinfl(satellites_num ~ width_cm + spine_cond +
            width_cm:spine_cond | width_cm + spine_cond,
          dist = "negbin", data=data)
```

## Modeling - Negative Binomial
\vskip1em
\textbf{Example:} \vskip.5em
\scriptsize
```{r}
summary(m2)[1]
```
