---
title: "Assignment 1"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

```{r setup}	
library(tidyverse)
library(ssstats)
```

Corporate is hosting a "Dunder Mifflin Sales Summit," and Michael Scott insists Scranton should look elite. HQ collects a one-time snapshot of performance and workstyle metrics for a large group of sales employees across multiple branches (Scranton, Stamford, Utica, Albany, etc.). Each row is one employee, measured during the same month (i.e., this is independent data). In the <u><b>week1.csv</b></u> data file, you will find the following variables:

- *annual_sales_usd*: annual sales for the employee (USD), annualized from consistent internal accounting rules,
- *client_calls_week*: average number of client calls per week,
- *meetings_week*: average number of client meetings per week,
- *emails_week*: average number of client-facing emails per week,
- *avg_call_minutes*: average call length (minutes),
- *lead_quality*: portfolio lead quality score (0–100),
- *crm_minutes_day*: average minutes/day spent in CRM,
- *coffee_cups_day*: average cups of coffee per day,
- *distractions_minutes_day*: average minutes/day lost to distractions (pranks, "meetings," etc.),
- *morale_score*: morale score (0–10) from anonymous internal survey, and
- *printer_jams_week*: average printer jams per week (yes, they tracked it).

**Data Management**

Download the data from Canvas and import it to R.

```{r}

```

**Model 1** (`m1`)

Fit a multiple linear regression model predicting annual sales (*annual_sales_usd*) as a function of the number of client calls per week (*client_calls_week*), the number of client meetings per week (*meetings_week*), and the number of client-facing emails per week (*emails_week*). Call this model `m1`.

```{r}

```

<!-- no actual response is needed for the modeling portion here, just your R output -->

1a. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} = \beta_0 + \beta_1 x_1 + ... \beta_k x_k
$$

1b. Is this a significant regression line? You must provide statistical justification for your answer.

```{r}

```

<!-- pretend that I do not know how to read or interpret
     R output; you must give a verbal response here -- I
     want you to tell me what the output means in context -->
     
1c. Fill in the following table:

| Predictor          | $\hat{\beta}_i$ (95% CI for $\beta_i$) | *p*-value |
|--------------------|:-----------------------------------------:|-----------:|
| client calls       |                                         |           |
| client meetings    |                                         |           |
| client emails      |                                         |           |

```{r}

```

1d. Which, if any, are significant predictors of annual sales? You must provide statistical justification for your answers.

```{r}

```

<!-- pretend that I do not know how to read or interpret
     R output; you must give a verbal response here -- I
     want you to tell me what the output means in context -->


**Model 2** (`m2`)

Fit a multiple linear regression model predicting annual sales (*annual_sales_usd*) as a function of the number of client calls per week (*client_calls_week*), the number of client meetings per week (*meetings_week*), the number of client-facing emails per week (*emails_week*), the number of cups of coffee per day (*coffee_cups_day*), and the time lost to distractions per day (*distractions_minutes_day*). Call this model `m2`.

```{r}

```

<!-- no actual response is needed for the modeling portion here, just your R output -->

2a. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} = \beta_0 + \beta_1 x_1 + ... \beta_k x_k
$$

2b. Is this a significant regression line? You must provide statistical justification for your answer.  Assume $\alpha=0.05$.

```{r}

```

<!-- pretend that I do not know how to read or interpret
     R output; you must give a verbal response here -- I
     want you to tell me what the output means in context -->
     
2c. Fill in the following table:

| Predictor          | $\hat{\beta}_i$ (95% CI for $\beta_i$) | *p*-value |
|--------------------|:---------------------------------------:|-----------:|
| client calls       |                                         |           |
| client meetings    |                                         |           |
| client emails      |                                         |           |
| coffee cups       |                                         |           |
| distractions      |                                         |           |

```{r}

```

2d. Which, if any, are significant predictors of annual sales? You must provide statistical justification for your answers. Assume $\alpha=0.05$.

```{r}

```

<!-- pretend that I do not know how to read or interpret
     R output; you must give a verbal response here -- I
     want you to tell me what the output means in context -->

**Comparing the Models**

3a. Let's compare the two models side-by-side. Fill in the following table:

<!-- please copy and paste from your tables above and pick out what you need -->

| Predictor | `m1` $\hat{\beta}_i$ | `m1` *p*-value | `m2` $\hat{\beta}_i$ | `m2` *p*-value |
|--------------------|:-----------:|-----------:|:-----------:|-----------:|
| client calls       |         |      |             |            |
| client meetings    |         |      |             |            |
| client emails      |         |      |             |            |
| coffee cups        | xxxxxxx | xxxx |             |            |
| distractions       | xxxxxxx | xxxx |             |            |

3b. How different are the model results? You should compare the coefficients and significance of each predictor across the two models.

<!-- pretend that I do not know how to read or interpret
     R output; you must give a verbal response here -- I
     want you to tell me what the output means in context -->

**Model Selection**

4a. Find the $R_{\text{adj}}^2$ for each model. What do these values mean? 

```{r}

```

* `m1`:

* `m2`:

<!-- remember to intepret the values for me :) -->

4b. Based on $R_{\text{adj}}^2$, which model is preferred? Why?

<!-- remember to justify your answer -->

4c. Find the AIC and BIC values for each model. 

```{r}

```

| Model | AIC  | BIC  |
|:-------:|:------:|:------:|
| `m1`  |      |      |
| `m2`  |      |      |

4d. Based on AIC and BIC, which model is preferred? Why?

<!-- remember to justify your answer -->

4e. Considering your answers to 4b and 4d, which model will you choose as your final model? Why?

<!-- remember to justify your answer -->

**Reporting the Final Model**

5a. Using the model chosen in 4e, paste in either the appropriate model summary table (from 1c or 2c).

<!-- literally -- scroll up, copy the appropriate table, scroll back here, then paste the table :) -->



5b. Provide formal interpretations for each predictor in the final model. Note that all predictors are shown for completeness; please remove any predictors that do not appear in your final model.

* client calls: <!-- your interpretation here --> 
* client meetings: <!-- your interpretation here -->
* client emails: <!-- your interpretation here -->
* coffee cups: <!-- your interpretation here -->
* distractions: <!-- your interpretation here -->


5c. Construct an appropriate data visualization to help explain the relationships in your final model. While there is no single "right" answer here, you should consider your audience (...potentially Michael Scott...) and what would be most helpful for them to understand what the math is suggesting. For uniformity (for grading purposes), let's have annual sales (*annual_sales_usd*) on the *y*-axis and number of client meetings (*meetings_week*) on the *x*-axis. For predicted values, please plug in the `median()` value for all other predictors in the model.

```{r}

```

5d. Pretend that you are preparing to meet with Michael Scott to review the analysis results. Summarize the key takeaways from your model and visualization. Remember your audience here -- [Michael is not a statistician](https://www.youtube.com/watch?v=IBJJrZ5LAVQ), so you should avoid jargon and technical language. What would be most helpful for him to know to improve sales performance at Dunder Mifflin?

