---
title: "Assignment 2"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    toc: true
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

```{r setup}
library(tidyverse)
library(ssstats)
library(broom)
```

After the success of Michael's "Sales Summit Scorecard," corporate decides to expand the analysis to understand structural differences across employees. This time, they collect a one-time snapshot of sales employees across multiple Dunder Mifflin branches, focusing on how role structure and branch culture relate to sales performance.

Each row in the dataset represents one employee. All variables summarize employee behavior and performance over the same time period. In the <u><b>week2.csv</b></u> data file (yes, we are reusing last week's data), you will find the following variables:

- *annual_sales_usd*: annual sales for the employee (USD),
- *sales_role*: sales role (Inside vs. Outside),
- *branch*: branch location (Scranton, Stamford, Utica, Albany),
- *client_calls_week*: average number of client calls per week,
- *meetings_week*: average number of client meetings per week,
- *emails_week*: average number of client-facing emails per week,
- *avg_call_minutes*: average call length (minutes),
- *lead_quality*: portfolio lead quality score (0–100),
- *crm_minutes_day*: average minutes/day spent in CRM,
- *coffee_cups_day*: average cups of coffee per day,
- *distractions_minutes_day*: average minutes/day lost to distractions (pranks, "meetings," etc.), and
- *morale_score*: morale score (0–10) from anonymous internal survey.


## Data Management

0a. Download the data from Canvas and import it to R OR import it from GitHub.

```{r}

```

0b. Examine what is contained in the *branch* variable by producing a frequency table.

```{r}

```

<!-- 
if you do not know how to find summary statistics in R, please see this lecture 
from my Biostatistics course: 
https://samanthaseals.github.io/STA4173Fa25/files/slides/01-2-estimation-in-R.html
-->

0c. Find summary statistics (mean and sd) for the continuous predictor variables in the dataset. Split this by *branch*.

```{r}

```

## Constructing the Model

1a. Fit a multiple linear regression model predicting annual sales (*annual_sales_usd*) as a function of the number of client calls per week (*client_calls_week*), the number of client meetings per week (*meetings_week*), and the branch (*branch*; use Scranton as the reference group). Call this model `m1`.

```{r}

```

1b. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

1c. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->
     
1d. Fill in the following table:

| Predictor          | $\hat{\beta}_i$ (95% CI for $\beta_i$) | *p*-value |
|--------------------|:-----------------------------------------:|-----------:|
| client calls       |                                         |           |
| client meetings    |                                         |           |
| branch (Stamford)  |                                         |           |
| branch (Utica)     |                                         |           |
| branch (Albany)    |                                         |           |

```{r}

```

1e. Which, if any, are significant predictors of annual sales? You must provide statistical justification for your answers. Assume $\alpha=0.05$.

```{r}

```

*Client Calls*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Client Meetings*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Branch*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

1f. Which branches, if any, are significantly different from Scranton? You must provide statistical justification for your answers. Assume $\alpha=0.05$.

```{r}

```

*Albany*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Stamford*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Utica*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

## Model Assumptions

2a. State the assumptions on the linear model.



2b. Construct the appropriate graph(s) to check assumptions listed in 2a.

```{r}

```

<!-- make sure that you follow only the methods outlined in lecture -->

2c. Do we meet the assumptions for the linear model? Why or why not? Does this mean our inference is valid or invalid?


## Model Interpretation

3a. Provide formal interpretations for each predictor in the model. 

- client calls: <!-- your interpretation here --> 
- client meetings: <!-- your interpretation here -->
- branch (Stamford): <!-- your interpretation here -->
- branch (Utica): <!-- your interpretation here -->
- branch (Albany): <!-- your interpretation here -->

3b. Find the adjusted $R^2$ for the model. What does this mean in the problem's context?

```{r}

```

3c. Construct an appropriate data visualization to help explain the relationships in your final model. While there is no single "right" answer here, you should consider your audience (...potentially Michael Scott...) and what would be most helpful for them to understand what the math is suggesting. As the outcome, annual sales (*annual_sales_usd*) should be on the *y*-axis. Pick the continuous predictor you beleive has the strongest relationship with annual sales as the *x*-axis. There should be a predicted line for every branch. To then construct predicted values, please plug in the `median()` value of all other continuous predictors.

```{r}

```

3d. Pretend that you are preparing to meet with Michael Scott to review the analysis results. Remember your audience here -- [Michael is not a statistician](https://www.youtube.com/watch?v=IBJJrZ5LAVQ), so you should avoid jargon and technical language. Write a brief paragraph to go with your visualization that explains the analysis results in plain language and include a suggestion for increasing annual sales. 














