---
title: "Assignment 2"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

After the success of Michael's "Sales Summit Scorecard," corporate decides to expand the analysis to understand structural differences across employees. This time, they collect a one-time snapshot of sales employees across multiple Dunder Mifflin branches, focusing on how role structure and branch culture relate to sales performance.

Each row in the dataset represents one employee. All variables summarize employee behavior and performance over the same time period. In the <u><b>week2.csv</b></u> data file, you will find the following variables:

- *annual_sales_usd*: annual sales for the employee (USD),
- *sale_role*: sales role (Inside vs. Outside),
- *branch*: branch location (Scranton, Stamford, Utica, Albany),
- *client_calls_week*: average number of client calls per week,
- *meetings_week*: average number of client meetings per week,
- *emails_week*: average number of client-facing emails per week,
- *avg_call_minutes*: average call length (minutes),
- *lead_quality*: portfolio lead quality score (0–100),
- *crm_minutes_day*: average minutes/day spent in CRM,
- *coffee_cups_day*: average cups of coffee per day,
- *distractions_minutes_day*: average minutes/day lost to distractions (pranks, "meetings," etc.), and
- *morale_score*: morale score (0–10) from anonymous internal survey.


**Data Management**

0a. Download the data from Canvas and import it to R.

```{r}

```

0b. Examine what is contained in the *sales_role* variable by producing a frequency table.

```{r}

```

0c. Examine what is contained in the *branch* variable by producing a frequency table.

```{r}

```

0d. Find summary statistics (mean and sd) for the continuous predictor variables in the dataset. Split this by *branch*.

```{r}

```

**Constructing the Model**

1a. Fit a multiple linear regression model predicting annual sales (*annual_sales_usd*) as a function of the number of client calls per week (*client_calls_week*), the number of client meetings per week (*meetings_week*), and the branch (*branch*; use Scranton as the reference group). Call this model `m1`.

```{r}

```

1b. Based on your output, state the resulting model. 

$$
\hat{y} = \beta_0 + \beta_1 x_1 + ... \beta_k x_k
$$

1c. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```
     
1d. Fill in the following table:

| Predictor          | $\hat{\beta}_i$ (95% CI for $\beta_i$) | *p*-value |
|--------------------|:-----------------------------------------:|-----------:|
| client calls       |                                         |           |
| client meetings    |                                         |           |
| branch (Stamford)  |                                         |           |
| branch (Utica)     |                                         |           |
| branch (Albany)    |                                         |           |

```{r}

```

1e. Which, if any, are significant predictors of annual sales? You must provide statistical justification for your answers. Assume $\alpha=0.05$.

```{r}

```

**Model Assumptions**

2a. State the assumptions on the linear model.



2b. Construct the appropriate graph(s) to check assumptions listed in 2a.

```{r}

```

2c. Do we meet the assumptions for the linear model? Are we okay to proceed with statistical inference?



**Model Interpretation**

3a. Provide formal interpretations for each predictor in the model. 

* client calls: <!-- your interpretation here --> 
* client meetings: <!-- your interpretation here -->
* branch (Stamford): <!-- your interpretation here -->
* branch (Utica): <!-- your interpretation here -->
* branch (Albany): <!-- your interpretation here -->

3b. Find the adjusted $R^2$ for the model. What does this mean in context?



3c. Construct an appropriate data visualization to help explain the relationships in your final model.For uniformity (for grading purposes), let's again have annual sales (*annual_sales_usd*) on the *y*-axis, the number of client meetings (*meetings_week*) on the *x*-axis, and the lines defined by the branch (*branch*). For predicted values, please plug plausible values for the other predictors in the model.

```{r}

```

3d. Pretend that you are preparing to meet with Michael Scott to review the analysis results. Summarize the key takeaways from your model and visualization. Remember your audience here -- [Michael is not a statistician](https://www.youtube.com/watch?v=IBJJrZ5LAVQ), so you should avoid jargon and technical language. What would be most helpful for him to know to improve sales performance at Dunder Mifflin? What has changed from last week's analysis?














