---
title: "Assignment 3"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

After the success of Michael's "Sales Summit Scorecard," corporate decides to expand the analysis to understand structural differences across employees. This time, they collect a one-time snapshot of sales employees across multiple Dunder Mifflin branches, focusing on how role structure and branch culture relate to sales performance.

Each row in the dataset represents one employee. All variables summarize employee behavior and performance over the same time period. In the <u><b>week2.csv</b></u> data file, you will find the following variables:

- *annual_sales_usd*: annual sales for the employee (USD),
- *sales_role*: sales role (Inside vs. Outside),
- *branch*: branch location (Scranton, Stamford, Utica, Albany),
- *client_calls_week*: average number of client calls per week,
- *meetings_week*: average number of client meetings per week,
- *emails_week*: average number of client-facing emails per week,
- *avg_call_minutes*: average call length (minutes),
- *lead_quality*: portfolio lead quality score (0–100),
- *crm_minutes_day*: average minutes/day spent in CRM,
- *coffee_cups_day*: average cups of coffee per day,
- *distractions_minutes_day*: average minutes/day lost to distractions (pranks, "meetings," etc.), and
- *morale_score*: morale score (0–10) from anonymous internal survey.


**Data Management**

0a. Download the data from Canvas and import it to R OR import it from GitHub.

```{r}

```

0b. Find summary statistics (mean and sd) for the continuous predictor variables in the dataset. Split this by *branch*.

<!-- yes, please reuse your code from last week -->

```{r}

```



**Constructing the Models**

1a. Fit a multiple linear regression model predicting annual sales (*annual_sales_usd*) as a function of client meetings per week (*meetings_week*), morale score (*morale_score*), and branch (*branch*; use Scranton as the reference group). Call this model `m1`.

```{r}

```

1b. Fit a multiple linear regression model predicting annual sales (*annual_sales_usd*) as a function of client meetings per week (*meetings_week*), morale score (*morale_score*), branch (*branch*; use Scranton as the reference group), and the interaction between client meetings per week and branch. Call this model `m2`.

```{r}

```

1c. Fit a multiple linear regression model predicting annual sales (*annual_sales_usd*) as a function of client meetings per week (*meetings_week*), morale score (*morale_score*), branch (*branch*; use Scranton as the reference group), the interaction between client meetings per week and branch, and the interaction between morale score and branch. Call this model `m3`.

```{r}

```

**Choosing the Model**

2a. Find the appropriate model fit statistics for each of the three models you fit in 1a–1c. 

```{r}

```

2b. Based on the statistics in 2a, which model are you choosing to report as the best model? You must justify your answer statistically.

Students should choose to report `m2` based on AIC/BIC (lowest value) or $R^2$ (highest value). 

**Investigating the Model**

3a. Report your chosen model. 

$$
\hat{y} = \beta_0 + \beta_1 x_1 + ... \beta_k x_k
$$

3b. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```

3c. Is the interaction between client meetings and branch significant? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```
     
3d. Stratify the model by branch.

$$
\begin{align*}
\hat{\text{annual sales }} | \text{ Albany} &= \hat{\beta}_0 + ... \\
\hat{\text{annual sales }} | \text{ Scranton} &= \hat{\beta}_0 + ... \\
\hat{\text{annual sales }} | \text{ Stamford} &= \hat{\beta}_0 + ... \\
\hat{\text{annual sales }} | \text{ Utica} &= \hat{\beta}_0 + ... \\
\end{align*}
$$

3e. Based on your stratified models, which branch is performing the best?  Which branch is performing the worst? You must provide statistical justification for your answers. 



**Visualizing the Model**

4a. Construct a data visualization to help explain the relationships in your stratified models. For uniformity (for grading purposes), let's have annual sales (*annual_sales_usd*) on the *y*-axis, the number of client meetings (*meetings_week*) on the *x*-axis, and the lines defined by the branch (*branch*). Plug in the median moral score when constructing predicted values.

```{r}


```

**Interpreting the Model**

5a. Provide formal interpretations for:

- client meetings when branch = Albany:
- client meetings when branch = Scranton:
- client meetings when branch = Stamford:
- client meetings when branch = Utica:

5b. Provide a formal interpretation for the effect of morale score.

5c. In 5a, why are we reporting branch-specific effects?

5d. In 5b, why are we providing overall interpretation, and not branch-specific?

5e. Pretend that you are preparing to meet with Michael Scott to review the analysis results. Remember your audience here -- [Michael is not a statistician](https://www.youtube.com/watch?v=IBJJrZ5LAVQ), so you should avoid jargon and technical language. Write a brief paragraph to go with your visualization that explains the analysis results in plain language and include a suggestion for increasing annual sales. 
















