---
title: "Assignment 4"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    toc: true
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

```{r setup}
library(tidyverse)
library(ssstats)
library(broom)
```

Jim and Pam are finally planning their wedding! Dwight is trying to convince them to have the wedding and reception at Schrute Farms, but both Jim and Pam are skeptical: they are concerned both about the amount of time guests will wait for food and the amount of time guests will actually want to spend on the dance floor. Their wedding planner manages to find data on previous events held at Schrute Farms. The data that was collected is as follows,

- *food_wait_min*: average time (minutes) the group waited before reaching the buffet line
- *dance_floor_prop*: the proportion of the reception time that guests spent on the dance floor
- *guests*: number of guests at the event
- *temp_f*: the outdoor temperature when the group entered the buffet line
- *dwight_logistics*: indicates if Dwight was (1) or was not (0) responsible for logistics during that time

Because Dwight has a special talent for either making things significantly better or much, much worse, the wedding planner believes that the effect of group size and temperature might depend on whether Dwight was running logistics.

## Data Import

0a. Download the data from Canvas and import it to R.

```{r}

```

<!-- no actual response is needed here, just your R output -->

## Model 1 (`m1`)

Because Dwight has a special talent for either making things significantly better or much, much worse, the wedding planner believes that the effect of group size and temperature might depend on whether Dwight was running logistics. We want to fit a multiple regression model predicting the time the group waited for food (*food_wait_min*) as a function of the number of guests (*guests*), the outdoor temperature (*temp_f*), if Dwight was involved with logistics (*dwight_logistics*), the interaction between number of guests and Dwight's involvement, and the interaction between temperature and Dwight's involvement. 

1a. Construct the histogram for the time waited before reaching the buffet line (*food_wait_min*).

```{r}

```

<!-- no actual response is needed here, just your R output -->

1b. Based on the histogram results from 1a, what distribution do you feel is most appropriate for modeling the outcome variable? You must justify your choice.

<!-- hint: do not say the normal distribution -- revisit the topics from this week -->


1c. Find the minimum and maximum values of the outcome variable. 

```{r}

```

<!-- no actual response is needed here, just your R output -->

1d. Based on the results from 1c and distribution choice in 1b, perform any additional data management that needs to be done prior to modeling. Explain why this data management was or was not necessary.


1e. Fit the appropriate regression model based on your distribution choice in 1b, including the interaction terms described above. Call this model `m1`.

```{r}

```

<!-- no actual response is needed for the modeling portion here, just your R output -->

1f. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

<!-- imperative for full credit: remove the pieces of the left hand side that you do not need -->

1g. Is this a significant regression line? You must provide statistical justification for your answer.

```{r}

```

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

1h. Perform the appropriate hypothesis tests to determine if either the interaction between Dwight's involvement and the outdoor temperature or the interaction between Dwight's involvement and the number of guests is significant. You must provide statistical justification for your answers.

```{r}

```

*Dwight $\times$ Temperature*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Dwight $\times$ Guests*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->


1i. Regardless of significance, stratify the models by Dwight's involvement in logistics.

- When Dwight is involved (*dwight_logistics* = 1):

- When Dwight is not involved (*dwight_logistics* = 0):

1j. Using the stratified models from 1i, provide formal interpretations for the following slopes:

- Slope for number of guests

    - when Dwight is involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    - when Dwight is not involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    
- Slope for temperature

    - when Dwight is involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    - when Dwight is not involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    
1k. Construct a data visualization to help explain the relationships in your stratified models. For uniformity, let's have time waited for food (*food_wait_min*) on the *y*-axis, the outdoor temperature (*temp_f*) on the *x*-axis, and lines defined by when Dwight is involved in logistics. We will plug in the overall median value of guests (*guests*).

```{r}

```

<!-- no actual response is needed for the modeling portion here, just your R output -->

1l. Pretend that you are preparing to meet with the wedding planner to review the analysis results. Remember your audience here -- you should avoid jargon and technical language. Write a brief paragraph to go with your visualization that explains the analysis results in plain language and include a suggestion for the wedding logistics.

## Model 2 (`m2`)

Because Dwight has a special talent for either making things significantly better or much, much worse, the wedding planner believes that the effect of group size and temperature might depend on whether Dwight was running logistics. Let's now fit a multiple regression model predicting proportion of time guests spent dancing (*dance_floor_prop*) as a function of the number of guests (*guests*), the outdoor temperature (*temp_f*), if Dwight was involved with logistics (*dwight_logistics*), the interaction between number of guests and Dwight's involvement, and the interaction between temperature and Dwight's involvement.

2a. Construct the histogram for the proportion of time guests spent dancing (*dance_floor_prop*).

```{r}

```

<!-- no actual response is needed here, just your R output -->

2b. Based on the histogram results from 2a, what distribution do you feel is most appropriate for modeling the outcome variable? You must justify your choice.

<!-- hint: do not say the normal distribution -- revisit the topics from this week -->


2c. Find the minimum and maximum values of the outcome variable. 

```{r}

```

<!-- no actual response is needed here, just your R output -->

2d. Based on the results from 2c and distribution choice in 2b, perform any additional data management that needs to be done prior to modeling. Explain why this data management was or was not necessary.


2e. Fit the appropriate regression model based on your distribution choice in 2b, including the interaction terms described above. Call this model `m2`.

```{r}

```


<!-- no actual response is needed for the modeling portion here, just your R output -->

2f. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

<!-- imperative for full credit: remove the pieces of the left hand side that you do not need -->

2g. Is this a significant regression line? You must provide statistical justification for your answer.

```{r}

```

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

2h. Perform the appropriate hypothesis tests to determine if either the interaction between Dwight's involvement and the outdoor temperature or the interaction between Dwight's involvement and the number of guests is significant. You must provide statistical justification for your answers.

```{r}

```

*Dwight $\times$ Temperature*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Dwight $\times$ Guests*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->


2i. Regardless of significance, stratify the models by Dwight's involvement in logistics.

- When Dwight is involved (*dwight_logistics* = 1):


- When Dwight is not involved (*dwight_logistics* = 0):


2j. Using the stratified models from 2i, provide formal interpretations for the following slopes:

- Slope for number of guests

    - when Dwight is involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    - when Dwight is not involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    
- Slope for temperature

    - when Dwight is involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    - when Dwight is not involved: <!-- your answer sandwiched to the right --> <!-- your answer sandwiched to the left -->
    
2k. Construct a data visualization to help explain the relationships in your stratified models. For uniformity, let's have time waited for proportion of time guests spent dancing (*dance_floor_prop*) on the *y*-axis, the outdoor temperature (*temp_f*) on the *x*-axis, and lines defined by when Dwight is involved in logistics. We will plug in the overall median value of guests (*guests*).

```{r}

```

<!-- no actual response is needed for the modeling portion here, just your R output -->

2l. Pretend that you are preparing to meet with the wedding planner to review the analysis results. Remember your audience here -- you should avoid jargon and technical language. Write a brief paragraph to go with your visualization that explains the analysis results in plain language and include a suggestion for the wedding logistics.

## Reflection

This section is **required** for graduate students and **optional** for undergraduate students. This section can be written more technically (for Dr. Seals, not Michael Scott), but you should still be explicitly clear in your writing. Please write one paragraph (hard limit) answering each of the following prompts.

<!-- Friends, please note that the prompts below are intentionally open-ended. There is no single "right" answer here, but there are definitely wrong answers. I ask these questions to make you critically think about and synthesize the class material, so the important thing is that you provide a thoughtful, well-reasoned response that demonstrates an appropriate level of understanding of the material.   -->

3a. Reflect on the modeling decisions you made for Model 1. Why is the modeling approach you chose more appropriate than ordinary linear regression for that outcome? Please consider and address what assumption(s) of linear regression would be violated if you had used it instead. 



3b. Reflect on the modeling decisions you made for Model 2. Why is the modeling approach you chose more appropriate than ordinary linear regression for that outcome? Please consider and address what assumption(s) of linear regression would be violated if you had used it instead.



3c. In both models, you have the same set of predictors but different outcomes. Compare the results across Model 1 and Model 2. Are the predictors that were significant in one model also significant in the other? What does this pattern (or lack of one) suggest about how group size, temperature, and Dwight's supervision relate to the time guests wait for food and the time guests spend on the dance floor at Schrute Farms?



3d. Suppose you are preparing for your meeting with the wedding planner. Are there any limitations of this analysis? What additional data would you want to collect to further inform recommendations? Are there other analyses you would suggest? (i.e., what would be your next steps if you were moving forward with this analysis?)