---
title: "Assignment 5"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    toc: true
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

```{r setup}
library(tidyverse)
library(ssstats)
library(broom)
```

Dunder Mifflin's Scranton branch is heading into its annual review cycle, and Michael Scott has decided that this quarter's employee performance will be evaluated using a highly sophisticated process.

For each employee, the company has recorded several measures of daily work habits:

- *sales_call_per_day*: the average number of sales calls an employee makes
- *hours_at_desk*: the average number of hours spent at their desk each day
- *motivational_poster*: whether Michael has recently hung a new motivational poster near the employee’s workspace (1 = yes, 0 = no)

Management is interested in two different outcomes:

- *promotion_considered*: whether the employee is being considered for a promotion (1 = yes, 0 = no)
- *performance_review*: the employee's performance review rating, which can be "Outstanding", "Adequate", or "Disaster"

## Data Management

0a. Download the data from Canvas and import it to R.

```{r}

```

<!-- no actual response is needed here, just your R output -->

## Model 1 (`m1`)

Our goal is to fit a multiple regression model predicting the probability that an employee is considered for promotion (*promotion_considered*) as a function of number of sales calls (*sales_call_per_day*), number of hours at desk (*hours_at_desk*), and if Michael has recently hung a new motivational poster (*motivational_poster*; 1 = yes, 0 = no).

1a. Construct a frequency table for if the employee is being considered for a promotion (*promotion_considered*).

```{r}

```

<!-- no actual response is needed here, just your R output -->

1b. Based on the frequency table in 1a, what modeling approach do you feel is most appropriate for modeling the outcome variable? Why?


1c. Based on the results from 1a and distribution choice in 1b, perform any additional data management that needs to be done prior to modeling. Explain why this data management was or was not necessary.

<!-- if you do not need data management, you do not need this code chunk -->
```{r}

```

1d. Construct the model described in the Model 1 set up. Call this model `m1`.

```{r}

```

<!-- no actual response is needed for the modeling portion here, just your R output -->

1e. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

<!-- imperative for full credit: remove the pieces of the left hand side that you do not need -->

1f. Fill out the following table,

| Predictor          | $\hat{\text{OR}}_i$ (95% CI for $\text{OR}_i$)  | *p*-value |
|--------------------|:---------------------------------------:|----------:|
| Number of Sales Calls    |                                         |           |
| Time at Desk (hours)  |                                         |           |
| Motivational Poster (yes vs. no)  |                                         |           |


```{r}

```

1g. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->


1h. Which, if any, are significant predictors of being up for promotion? You must provide statistical justification for your answers. Assume $\alpha=0.05$.

```{r}

```

*Sales Calls*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Hours at Desk*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Motivational Poster*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

1i. Provide formal interpretations for each predictor in the model. 

- sales calls: <!-- your interpretation here -->
- hours at desk: <!-- your interpretation here -->
- motivational poster: <!-- your interpretation here -->

1j. Construct an appropriate data visualization to help explain the relationships in your model. For uniformity (for grading purposes), let's have promotion consideration (*promotion_considered*) on the *y*-axis, the number of hours at their desk (*hours_at_desk*) on the *x*-axis, and the lines defined by the presence of a motivational poster (*motivational_poster*). For predicted values, plug in plausible values for other predictors in the model.

```{r}

```

1k. Pretend that you are preparing to meet with Michael Scott to review the analysis results. Remember your audience here -- [Michael is not a statistician](https://www.youtube.com/watch?v=IBJJrZ5LAVQ), so you should avoid jargon and technical language. Write a brief paragraph to go with your visualization that explains the analysis results in plain language.



## Model 2 (`m2`)

Our goal is now to fit a multiple regression model predicting the result of an employee's evaluation (*performance_review*) as a function of number of sales calls (*sales_call_per_day*), number of hours at desk (*hours_at_desk*), and if Michael has recently hung a new motivational poster (*motivational_poster*; 1 = yes, 0 = no). 

2a. Construct the frequency table for the second outcome, the employee's performance review rating (*performance_review*).

```{r}

```

<!-- no actual response is needed here, just your R output -->

2b. Based on the frequency table results from 2a what modeling approach do you feel is most appropriate for modeling the outcome variable?


2c. Based on the results from 2a and distribution choice in 2b, perform any additional data management that needs to be done prior to modeling. Explain why this data management was or was not necessary.

<!-- if you do not need data management, you do not need this code chunk -->
```{r}

```

2d. Construct the model described in the Model 2 set up. Call this model `m2`.

```{r}

```

<!-- no actual response is needed for the modeling portion here, just your R output -->

2e. Based on your output, state the resulting models. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\begin{align*}
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} &= \beta_0 + \beta_1 x_1 + ... + \beta_k x_k \\
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} &= \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
\end{align*}
$$

<!-- imperative for full credit: remove the pieces of the left hand side that you do not need -->


2f. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

2g. Which, if any, are overall significant predictors of evaluation result? You must provide statistical justification for your answers. Assume $\alpha=0.05$.

```{r}

```

*Sales Calls*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Hours at Desk*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

*Motivational Poster*

**Hypotheses**:

- $H_0:$
- $H_1:$

**Test Statistic and $p$-value**

- $statistic =$ <!-- please replace the word "statistic" with the type of test statistic you have (e.g., F, t, etc.) -->
- $p=$

**Rejection Region**

- Reject $H_0$ if ...

**Conclusion and Interpretation**

- Reject or fail to reject $H_0$. <!-- pick the appropriate one... do not leave as is -->
- There is or is not sufficient evidence to conclude... <!-- pick the appropriate choice & write the conclusion in the problem's context -->

2h. Pick <u>one</u> of the models reported in 2e and provide formal interpretations for each predictor in the model.

- sales calls: <!-- your interpretation here -->
- hours at desk: <!-- your interpretation here -->
- motivational poster: <!-- your interpretation here -->

2i. Construct an appropriate data visualization to help explain the relationships in your model. For uniformity (for grading purposes), let's have performance review (*performance_review*) on the *y*-axis, the number of hours at their desk (*hours_at_desk*) on the *x*-axis, and the lines defined by the presence of a motivational poster (*motivational_poster*). For predicted values, plug in plausible values for other predictors in the model.

```{r}

```

2j. Pretend that you are preparing to meet with Michael Scott to review the analysis results. Remember your audience here -- [Michael is not a statistician](https://www.youtube.com/watch?v=IBJJrZ5LAVQ), so you should avoid jargon and technical language. Write a brief paragraph to go with your visualization that explains the analysis results in plain language.

## Reflection

This section is **required** for graduate students and **optional** for undergraduate students. This section can be written more technically (for Dr. Seals, not Michael Scott), but you should still be explicitly clear in your writing. Please write one paragraph (hard limit) answering each of the following prompts.

<!-- Friends, please note that the prompts below are intentionally open-ended. There is no single "right" answer here, but there are definitely wrong answers. I ask these questions to make you critically think about and synthesize the class material, so the important thing is that you provide a thoughtful, well-reasoned response that demonstrates an appropriate level of understanding of the material.   -->

3a. Reflect on the modeling decisions you made for Model 1. Why is the modeling approach you chose more appropriate than ordinary linear regression for that outcome? Please consider and address what assumption(s) of linear regression would be violated if you had used it instead.

3b. Reflect on the modeling decisions you made for Model 2. Why is the modeling approach you chose more appropriate than ordinary linear regression for that outcome? Please consider and address what assumption(s) of linear regression would be violated if you had used it instead.

3c. Both models use the same set of predictors but different outcomes. Compare the results across Model 1 and Model 2. Are the predictors that were significant in one model also significant in the other? What does this suggest, if anything, about the relationship between being considered for promotion and performance review ratings?

3d. Suppose you are preparing for your meeting with the higher ups at Dunder Mifflin. Are there any limitations of this analysis? What additional data would you want to collect to further inform recommendations? Are there other analyses you would suggest? (i.e., what would be your next steps if you were moving forward with this analysis?)