---
title: "Assignment 6"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

```{r setup}
library(tidyverse)
library(ssstats)
library(broom)
```

Dunder Mifflin's Scranton branch is piloting a new "productivity initiative" designed by Michael Scott. Corporate has asked the analytics team to evaluate whether the initiative is improving performance or just creating more chaos.

You are given a dataset containing information about employees across several workdays. For each employee-day observation, the dataset includes:

- *meetings*: number of meetings attended that day
- *coffee_cups*: number of cups of coffee consumed
- *dwight_supervision*: whether Dwight directly supervised the employee that day 
- *sales_calls*: number of successful sales calls completed in a day
- *incidents*: number of unusual workplace incidents observed per day

0a. Download the data from Canvas and import it to R.

```{r}

```

<!-- no actual response is needed here, just your R output -->

**Model 1**

1a. We are interested in modeling the number of incidents observed in a day. Based on the distribution of this outcome, which modeling approach would be most appropriate? Justify your answer.

```{r}

```

<!-- your answer below here -->

<!-- your answer above here -->


1b. Fit an appropriate model to predict the number of incidents based on the number of meetings, cups of coffee, and whether Dwight supervised the employee.  Call this model `m1`.

```{r}

```

<!-- no actual response is needed here, just your R output -->

1c. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

1d. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```


<!-- your answer below here -->

<!-- your answer above here -->

1e. Which, if any, are significant predictors of the number of incidents observed? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```


<!-- your answer below here -->

<!-- your answer above here -->

1f. Interpret the following effects:

- **number of meetings**: 
    
- **cups of coffee**:
    
- **Dwight's supervision**: 
    
1g. Construct a basic visualization to help you explain model results. Clearly state the following:

- What is on the *x*-axis?
    
- What is on the *y*-axis?
    
- What will the line(s) represent?

```{r}

```

**Model 2**

2a. We are interested in modeling the number of successful calls completed in a day. Based on the distribution of this outcome, which modeling approach would be most appropriate? Justify your answer.

```{r}

```

<!-- your answer below here -->

<!-- your answer above here -->

2b. Fit an appropriate model to predict the number of successful calls completed in a day based on the number of meetings, cups of coffee, and whether Dwight supervised the employee.  Call this model `m2`.

```{r}

```

<!-- no actual response is needed here, just your R output -->

2c. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

2d. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```


<!-- your answer below here -->

<!-- your answer above here -->


2e. Which, if any, are significant predictors of the number of incidents observed? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```

<!-- your answer below here -->

<!-- your answer above here -->

2f. Interpret the following effects:

- **number of meetings**: 
    
- **cups of coffee**:
    
- **Dwight's supervision**: 
    
2g. Construct a basic visualization to help you explain model results. Clearly state the following:

- What is on the *x*-axis?
    
- What is on the *y*-axis?
    
- What will the line(s) represent?

```{r}

```    

**Summary**

3. Pick either `m1` or `m2` and summarize your findings for Michael. Be sure to include interpretations of the predictors, significance of the model, and any visualizations you created.

<!-- your answer below here -->

<!-- your answer above here -->

**Exploration Time!**

4a. Re-fit `m1` using the normal distribution and examine relevant statistical inference. How different are the results?

```{r}

```

<!-- your answer below here -->

<!-- your answer above here -->

4b. Check the assumptions on the normal distribution. Remember to draw a conclusion.

```{r}

```

<!-- your answer below here -->

<!-- your answer above here -->

4c. Re-fit `m2` using the normal distribution and examine relevant statistical inference. How different are the results?

```{r}

```

<!-- your answer below here -->

<!-- your answer above here -->

4d. Check the assumptions on the normal distribution. Remember to draw a conclusion.

```{r}

```

<!-- your answer below here -->

<!-- your answer above here -->

**Research Time!** 

This section is a **bonus** section for undergraduate students, but a **required** section for graduate students.
     
5a. What are the zero-inflated Poisson and zero-inflated negative binomial models? When are they appropriate to employ?

<!-- your answer below here -->

<!-- your answer above here -->

<!-- unless you have extensive experience in statistical modeling, this question cannot be answered without any research :) zero credit will be awarded for answers without proper citations -->

5b. Why aren't the zero-inflated models appropriate for this dataset? What would have to be different about the data for these models to be appropriate?

<!-- your answer below here -->

<!-- your answer above here -->