---
title: "Assignment 6"
author: "Your Name Here"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    df-print: paged
    toc: true
    self-contained: true
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

```{r setup}
library(tidyverse)
library(ssstats)
library(broom)
```

Dunder Mifflin's Scranton branch is piloting a new "productivity initiative" designed by Michael Scott. Corporate has asked the analytics team to evaluate whether the initiative is improving performance or just creating more chaos.

You are given a dataset containing information about employees across several workdays. For each employee-day observation, the dataset includes:

- *meetings*: number of meetings attended that day
- *coffee_cups*: number of cups of coffee consumed
- *dwight_supervision*: whether Dwight directly supervised the employee that day 
- *sales_calls*: number of successful sales calls completed in a day
- *incidents*: number of unusual workplace incidents observed per day

## Data Management and Exploration

0a. Download the data from Canvas and import it to R.

```{r}

```

<!-- no actual response is needed here, just your R output -->

0b. Construct the histogram for the number of unusual workplace incidents observed per day (*incidents*).

```{r}

```

<!-- no actual response is needed here, just your R output -->

0c. Based on the histogram results from 0b, what distribution do you feel is most appropriate for modeling the number of unusual workplace incidents observed per day?

0d. Construct the histogram for the number of successful calls completed in a day (*sales_calls*).

```{r}

```

<!-- no actual response is needed here, just your R output -->

0e. Based on the histogram results from 0d, what distribution do you feel is most appropriate for modeling the number of successful calls completed in a day?



## Model 1 (`m1`)

1a. Fit an appropriate model to predict the number of incidents based on the number of meetings, cups of coffee, and whether Dwight supervised the employee.  Call this model `m1`.

```{r}

```

<!-- no actual response is needed here, just your R output -->

1b. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

1c. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```


1d. Which, if any, are significant predictors of the number of incidents observed? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```


1e. Interpret the following effects:

- number of meetings: 
    
- cups of coffee:
    
- Dwight's supervision: 
    
1f. Construct a basic visualization to help you explain model results. Clearly state the following:

- What is on the *x*-axis?
    
- What is on the *y*-axis?
    
- What will the line(s) represent?

```{r}

```

## Model 2 (`m2`)

2a. Fit an appropriate model to predict the number of successful calls completed in a day based on the number of meetings, cups of coffee, and whether Dwight supervised the employee.  Call this model `m2`.

```{r}

```

<!-- no actual response is needed here, just your R output -->

2b. Based on your output, state the resulting model. 

<!-- below is the basic LaTeX for formatting a regression line; please update it with numbers from your model based on the results above -->

$$
\hat{y} \text{ OR } \hat{\ln(y)} \text{ OR } \hat{\text{logit}(y)} = \beta_0 + \beta_1 x_1 + ... + \beta_k x_k
$$

2c. Is this a significant regression line? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```


2d. Which, if any, are significant predictors of the number of successful calls completed in a day? You must provide statistical justification for your answer. Assume $\alpha=0.05$.

```{r}

```

2e. Interpret the following effects:

- number of meetings: 
    
- cups of coffee:
    
- Dwight's supervision: 
    
2f. Construct a basic visualization to help you explain model results. Clearly state the following:

- What is on the *x*-axis?
    
- What is on the *y*-axis?
    
- What will the line(s) represent?

```{r}

```    

## Summary

3. Pick either `m1` or `m2` and summarize your findings for Michael. Be sure to include interpretations of the predictors, significance of the model, and any visualizations you created.

## Exploration Time!

4a. Re-fit `m1` using the normal distribution and examine relevant statistical inference. How different are the results?

```{r}

```


4b. Check the assumptions on the normal distribution. Remember to draw a conclusion.

```{r}

```


4c. Re-fit `m2` using the normal distribution and examine relevant statistical inference. How different are the results?

```{r}

```

4d. Check the assumptions on the normal distribution. Remember to draw a conclusion.

```{r}

```

## Research Time!

This section is a **bonus** section for undergraduate students, but a **required** section for graduate students.
     
5a. What are the zero-inflated Poisson and zero-inflated negative binomial models? When are they appropriate to employ?

<!-- unless you have extensive experience in statistical modeling, this question cannot be answered without any research :) zero credit will be awarded for answers without proper citations -->

5b. Why aren't the zero-inflated models appropriate for this dataset? What would have to be different about the data for these models to be appropriate?
